### 2023-07-20 paper.Rmd
Now using .Rmd file.

---
title: 'MR-PFU: A suite of R packages for energy conversion chain analysis'
tags:
- energy
- exergy
- R
- matrix mathematics
- energy efficiency
authors:
- name: Matthew Kuperus Heun
  orcid: 0000-0002-7438-214X
  affiliation: 1
- name: Zeke Marshall
  orcid: 0000-0001-9260-7827
  affiliation: 2
- name: Paul E. Brockway
  orcid: 0000-0001-6925-8040
  affiliation: 2
- name: Emmanuel Aramendia
  orcid: 0000-0002-5964-6776
  affiliation: 2
affiliations:
- name: Engineering Department, Calvin University
  index: 1
- name: School of Earth and Environment, Leeds University
  index: 2
date: "23 May 2023"
bibliography: paper.bib
output:
  html_document:
    keep_md: yes
---

Max length: 250--1000 words

# Summary

Energy flows through society enable economic activity and facilitate human flourishing.
To understand economic growth and human well-being, 
the field of energy analysis (SEA) evaluates energy flows
from the primary stage 
(coal, oil, natural gas) 
to the final stage
(energy purchased by consumers, such as refined petroleum and electricity)
and ultimately to the useful stage
(heat, mechanical drive, and light),
the last energy before energy services
(such as transport, thermal comfort, and illumination).
Societal exergy analysis (SEA), a variant of energy analysis,
quantifies energy flows as exergy, 
the physical work potential of energy.

The forces on stars, galaxies, and dark matter under external gravitational
fields lead to the dynamical evolution of structures in the universe. The orbits
of these bodies are therefore key to understanding the formation, history, and
future state of galaxies. The field of "galactic dynamics," which aims to model
the gravitating components of galaxies to study their structure and evolution,
is now well-established, commonly taught, and frequently used in astronomy.
Aside from toy problems and demonstrations, the majority of problems require
efficient numerical tools, many of which require the same base code (e.g., for
performing numerical orbit integration).

# Statement of need

`Gala` is an Astropy-affiliated Python package for galactic dynamics. Python
enables wrapping low-level languages (e.g., C) for speed without losing
flexibility or ease-of-use in the user-interface. The API for `Gala` was
designed to provide a class-based and user-friendly interface to fast (C or
Cython-optimized) implementations of common operations such as gravitational
potential and force evaluation, orbit integration, dynamical transformations,
and chaos indicators for nonlinear dynamics. `Gala` also relies heavily on and
interfaces well with the implementations of physical units and astronomical
coordinate systems in the `Astropy` package [@astropy] (`astropy.units` and
`astropy.coordinates`).

`Gala` was designed to be used by both astronomical researchers and by
students in courses on gravitational dynamics or astronomy. It has already been
used in a number of scientific publications [@Pearson:2017] and has also been
used in graduate courses on Galactic dynamics to, e.g., provide interactive
visualizations of textbook material [@Binney:2008]. The combination of speed,
design, and support for Astropy functionality in `Gala` will enable exciting
scientific explorations of forthcoming data releases from the *Gaia* mission
[@gaia] by students and experts alike.

# Example usage

```{r}
library(Recca)
```

# Mathematics

Single dollars ($) are required for inline mathematics e.g. $f(x) = e^{\pi/x}$

Double dollars make self-standing equations:

$$\Theta(x) = \left\{\begin{array}{l}
0\textrm{ if } x < 0\cr
1\textrm{ else}
\end{array}\right.$$

You can also use plain \LaTeX for equations
\begin{equation}\label{eq:fourier}
\hat f(\omega) = \int_{-\infty}^{\infty} f(x) e^{i\omega x} dx
\end{equation}
and refer to \autoref{eq:fourier} from text.

# Citations

Citations to entries in paper.bib should be in
[rMarkdown](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)
format.

If you want to cite a software repository URL (e.g. something on GitHub without a preferred
citation) then you can do it with the example BibTeX entry below for @fidgit.

For a quick reference, the following citation commands can be used:
- `@author:2001`  ->  "Author et al. (2001)"
- `[@author:2001]` -> "(Author et al., 2001)"
- `[@author1:2001; @author2:2001]` -> "(Author1 et al., 2001; Author2 et al., 2002)"

# Figures

Figures can be included like this:
![Caption for example figure.\label{fig:example}](figure.png)
and referenced from text using \autoref{fig:example}.

Figure sizes can be customized by adding an optional second parameter:
![Caption for example figure.](figure.png){ width=20% }

# Acknowledgements

We acknowledge contributions from Brigitta Sipocz, Syrtis Major, and Semyeong
Oh, and support from Kathryn Johnston during the genesis of this project.

# References

### 2023-07-20 paper.Rmd
Updated paper.

@@
 To understand economic growth and human well-being, 
-the field of energy analysis (SEA) evaluates energy flows
+the field of energy analysis evaluates energy flows
 from the primary stage 
@@
 the last energy before energy services
-(such as transport, thermal comfort, and illumination).
-Societal exergy analysis (SEA), a variant of energy analysis,
+(transport, thermal comfort, and illumination).
+Societal exergy analysis (SEA), an extension of energy analysis,
 quantifies energy flows as exergy, 

### 2023-07-20 paper.Rmd
Better line breaks.

@@
 Figures can be included like this:
+
 ![Caption for example figure.\label{fig:example}](figure.png)
@@
 Figure sizes can be customized by adding an optional second parameter:
+
 ![Caption for example figure.](figure.png){ width=20% }

### 2023-07-20 paper.Rmd
Working on references.

@@
   index: 2
-date: "23 May 2023"
+date: "20 July 2023"
 bibliography: paper.bib
@@
 the physical work potential of energy.
+Historically, SEA has analyzed energy flows through specific countries
+using Excel spreadsheets, 
+with data in varying and inconsistent formats 
+[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
 
@@
 \begin{equation}\label{eq:fourier}
-\hat f(\omega) = \int_{-\infty}^{\infty} f(x) e^{i\omega x} dx
+\hat{f}(\omega) = \int_{-\infty}^{\infty} f(x) e^{i\omega x} dx
 \end{equation}

### 2023-07-20 paper.Rmd
Move statement of need section.
Comment unneeded text.

@@
 the physical work potential of energy.
+
+<!-- The forces on stars, galaxies, and dark matter under external gravitational -->
+<!-- fields lead to the dynamical evolution of structures in the universe. The orbits -->
+<!-- of these bodies are therefore key to understanding the formation, history, and -->
+<!-- future state of galaxies. The field of "galactic dynamics," which aims to model -->
+<!-- the gravitating components of galaxies to study their structure and evolution, -->
+<!-- is now well-established, commonly taught, and frequently used in astronomy. -->
+<!-- Aside from toy problems and demonstrations, the majority of problems require -->
+<!-- efficient numerical tools, many of which require the same base code (e.g., for -->
+<!-- performing numerical orbit integration). -->
+
+# Statement of need
+
 Historically, SEA has analyzed energy flows through specific countries
@@
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
-
-The forces on stars, galaxies, and dark matter under external gravitational
-fields lead to the dynamical evolution of structures in the universe. The orbits
-of these bodies are therefore key to understanding the formation, history, and
-future state of galaxies. The field of "galactic dynamics," which aims to model
-the gravitating components of galaxies to study their structure and evolution,
-is now well-established, commonly taught, and frequently used in astronomy.
-Aside from toy problems and demonstrations, the majority of problems require
-efficient numerical tools, many of which require the same base code (e.g., for
-performing numerical orbit integration).
-
-# Statement of need
 

### 2023-07-21 paper.Rmd
Updated statement of need.

@@
 Historically, SEA has analyzed energy flows through specific countries
-using Excel spreadsheets, 
+using spreadsheets, 
 with data in varying and inconsistent formats 
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
+Scientists in the field wanted to expand to cover all countries, but
+the spreadsheet approach to data analysis would not scale.
+We developed a series of `R` packages to enable 
+analyses for all countries, as described in the following table.
 

### 2023-07-23 paper.Rmd
Added BibTeX version of database reference.

@@
 
-<!-- The forces on stars, galaxies, and dark matter under external gravitational -->
-<!-- fields lead to the dynamical evolution of structures in the universe. The orbits -->
-<!-- of these bodies are therefore key to understanding the formation, history, and -->
-<!-- future state of galaxies. The field of "galactic dynamics," which aims to model -->
-<!-- the gravitating components of galaxies to study their structure and evolution, -->
-<!-- is now well-established, commonly taught, and frequently used in astronomy. -->
-<!-- Aside from toy problems and demonstrations, the majority of problems require -->
-<!-- efficient numerical tools, many of which require the same base code (e.g., for -->
-<!-- performing numerical orbit integration). -->
-
 # Statement of need
@@
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
-Scientists in the field wanted to expand to cover all countries, but
+The authors and others in the field
+wanted to expand to cover all countries, but
 the spreadsheet approach to data analysis would not scale.
@@
 analyses for all countries, as described in the following table.
+
+Reference for data repository [@Marshall:2023aa].
 

### 2023-07-23 paper.Rmd
Working on paper and refs.

@@
 to the final stage
-(energy purchased by consumers, such as refined petroleum and electricity)
-and ultimately to the useful stage
+(energy purchased by consumers, such as refined petroleum and electricity),
+to the useful stage
 (heat, mechanical drive, and light),
-the last energy before energy services
+and sometimes to energy services
 (transport, thermal comfort, and illumination).
@@
 quantifies energy flows as exergy, 
-the physical work potential of energy.
+the mechanical work potential of energy.
 
@@
 
-Historically, SEA has analyzed energy flows through specific countries
+Historically, SEA has analyzed energy flows through individual countries
 using spreadsheets, 
@@
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
-The authors and others in the field
-wanted to expand to cover all countries, but
-the spreadsheet approach to data analysis would not scale.
-We developed a series of `R` packages to enable 
+The authors of this paper and others in the field
+wanted to expand SEA to cover all countries, but
+the spreadsheet approach to data analysis is not scalable.
+
+To achieve scalable societal energy and exergy analysis, 
+several sub-problems needed to be solved.
+The first problem is the format for data storage. 
+We use the Physical Supply-Use Table (PSUT) data structure,
+a matrix approach to describing energy flows from 
+extraction to
+processing stages and, ultimately, 
+to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
+A second problem is that `R` data fames.
+
+It would be convenient to perform matrix mathematics in `R` data frames
+as easily as scalar mathematics with the `dplyr` package from the `tidyverse`.
+To do that, a way to embed matrices in a cell of a data frame is required.
+
+Further, a means of manipulating row and column names
+
+One problem is da
+
+that different countries, 
+and different years for a given country,
+will have varying energy products and
+
+To solve these problems and 
+to meet the need for the SEA community,
+we developed a suite of `R` packages.
+
+# Design of `R` packages
+
+to enable 
 analyses for all countries, as described in the following table.
@@
 
-# Example usage
+# Example
 
@@
 ```
+
+# Conclusion
+
+# Acknowledgements
 

### 2023-07-23 paper.Rmd
Add a table.

@@
 
-Energy flows through society enable economic activity and facilitate human flourishing.
+Energy flows through societal energy conversion chains (ECCs)
+enable economic activity and facilitate human flourishing.
 To understand economic growth and human well-being, 
-the field of energy analysis evaluates energy flows
+the field of energy analysis evaluates ECCs
 from the primary stage 
@@
 Societal exergy analysis (SEA), an extension of energy analysis,
-quantifies energy flows as exergy, 
+quantifies ECCs as exergy, 
 the mechanical work potential of energy.
@@
 
-Historically, SEA has analyzed energy flows through individual countries
-using spreadsheets, 
+Historically, SEA has analyzed the ECCs of individual countries
+using a large number of linked spreadsheets, 
 with data in varying and inconsistent formats 
@@
 wanted to expand SEA to cover all countries, but
-the spreadsheet approach to data analysis is not scalable.
+the spreadsheet approach to data analysis was
+deemed not scalable.
+We created a suite of `R` packages to provide scalable SEA 
+for every country in the world.
+This paper describes the design of those packages and
+demonstrates briefly their use.
 
-To achieve scalable societal energy and exergy analysis, 
-several sub-problems needed to be solved.
-The first problem is the format for data storage. 
-We use the Physical Supply-Use Table (PSUT) data structure,
+# Design of `R` packages
+
+Several decisions influenced the design and APIs for the packages.
+The decision involves data format. 
+We selected the Physical Supply-Use Table (PSUT) data structure,
 a matrix approach to describing energy flows from 
@@
 to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
-A second problem is that `R` data fames.
+Given the decision to use PSUT matrices to store ECC data,
+packages were created to enable manipulation of matrices
+within the tidyverse, as shown in the following table.
+
+| Package | Function |
+|---------|----------|
+| [RCLabels](https://github.com/MatthewHeun/RCLabels) | Manipulate row and column matrices |
 
 It would be convenient to perform matrix mathematics in `R` data frames
-as easily as scalar mathematics with the `dplyr` package from the `tidyverse`.
+as easily as scalar mathematics in data frames.
+
+with the `dplyr` package from the `tidyverse`.
 To do that, a way to embed matrices in a cell of a data frame is required.
@@
 we developed a suite of `R` packages.
-
-# Design of `R` packages
 

### 2023-07-23 paper.Rmd
Matrix tables.

@@
 The decision involves data format. 
-We selected the Physical Supply-Use Table (PSUT) data structure,
+We selected the Physical Supply-Use Table (PSUT) framework,
 a matrix approach to describing energy flows from 
@@
 to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
+The PSUT framework succinctly describes the flow of energy carriers
+(products in PSUT terminology) 
+among energy conversion machines 
+(industries in PSUT terminology)
+in a set of four matrices as described in the following table.
+
+| Matrix | rows x columns     | Name            | Description                                              |
+|:-------|:-------------------|:----------------|:---------------------------------------------------------|
+| **R**  | industry x product | Resource matrix | Contains exogeneous energy inputs to an ECC              |
+| **U**  | product x industry | Use matrix      | Describes how each energy conversion device uses energy  |
+| **V**  | industry x product | Make matrix     | Describes how each energy conversion device makes energy |
+| **Y**  | product x industry | Final demand matrix | Describes how each energy carrier is consumed        |
+
+energy flows through various 
+The matrix approach to data storage brings the challenge that 
+different countries, 
+and different years for a given country,
+have varying energy products and varying energy processing machines.
+
 Given the decision to use PSUT matrices to store ECC data,
@@
 |---------|----------|
-| [RCLabels](https://github.com/MatthewHeun/RCLabels) | Manipulate row and column matrices |
+| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulate row and column matrices |
+| [matsbyname](https://matthewheun.github.io/matsbyname/)  | Perform matrix mathematics that respects row and column names |
 
@@
 
-that different countries, 
-and different years for a given country,
-will have varying energy products and
+that
 

### 2023-07-23 paper.Rmd
Added RUVY table.

--- 2023-07-23 paper.Rmd
@@
 to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
+The PSUT framework succinctly describes the flow of energy carriers
+(``products'' in PSUT terminology) 
+among energy conversion machines 
+(``industries'' in PSUT terminology)
+in a set of four matrices as described in the following table.
+
+| Matrix      | rows x columns     | Name                           | Description                                              |
+|:------------|:-------------------|:-------------------------------|:---------------------------------------------------------|
+| **R**       | industry x product | Resource matrix                | Contains exogeneous energy inputs to an ECC              |
+| **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy  |
+| **U_EIOU**  | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
+| **V**       | industry x product | Make matrix                    | Describes how each energy conversion device makes energy |
+| **Y**       | product x industry | Final demand matrix            | Describes how each energy carrier is consumed            |
+
+The matrix approach to data storage brings the challenge that 
+different countries, 
+and different years for a given country,
+have varying energy products and varying energy processing machines.
+
 Given the decision to use PSUT matrices to store ECC data,

### 2023-07-23 paper.Rmd
More on the paper.

@@
 and different years for a given country,
-have varying energy products and varying energy processing machines.
+have varying energy energy carriers (products) and
+varying energy conversion machines (industries).
+To get around this challenge, a new `R` package 
+[matsbyname](https://matthewheun.github.io/matsbyname/)
+enables matrix mathematics that respects row and column names, 
+inserting `0` rows or columns where needed.
 
-Given the decision to use PSUT matrices to store ECC data,
-packages were created to enable manipulation of matrices
-within the tidyverse, as shown in the following table.
+Furthermore, 
+it would be convenient to perform matrix mathematics in `R` data frames
+as easily as scalar mathematics in data frames with [tidyverse](https://www.tidyverse.org) syntax. 
+[matsindf](https://matthewheun.github.io/matsindf/) enables this functionality.
 
 | Package | Function |
-|---------|----------|
-| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulate row and column matrices |
+|:--------|:---------|
+| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulate row and column names in [matsindf](https://matthewheun.github.io/matsindf/) data frames |
 | [matsbyname](https://matthewheun.github.io/matsbyname/)  | Perform matrix mathematics that respects row and column names |
+| [matsindf](https://matthewheun.github.io/matsindf/)      | Store matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
 
-It would be convenient to perform matrix mathematics in `R` data frames
-as easily as scalar mathematics in data frames.
+The packages in the table above are applicable to many problem domains and
+are available on [CRAN](https://cran.r-project.org).
 
-with the `dplyr` package from the `tidyverse`.
-To do that, a way to embed matrices in a cell of a data frame is required.
+Additional packages specific to ECC analysis within SEA are available on GitHub
+and described in the following table.
 
-Further, a means of manipulating row and column names
-
-One problem is da
-
-that
-
-To solve these problems and 
-to meet the need for the SEA community,
-we developed a suite of `R` packages.
-
-to enable 
-analyses for all countries, as described in the following table.
+| Package | Function |
+|:--------|:---------|
+| [Recca](https://matthewheun.github.io/Recca/) | Functions to perform `R` energy conversion chain analysis |
 

### 2023-07-23 paper.Rmd
Working on the paper.

@@
 |:--------|:---------|
-| [Recca](https://matthewheun.github.io/Recca/) | Functions to perform `R` energy conversion chain analysis |
+| [IEATools](https://matthewheun.github.io/IEATools/) | Converts IEA data to PSUT format |
+| [Recca](https://matthewheun.github.io/Recca/)       | Performs `R` energy conversion chain analysis |
+| [PFUDatabase](https://energyeconomydecoupling.github.io/PFUDatabase/) | The [targets](https://books.ropensci.org/targets/) pipeline to create a data frame of PSUT matrices |
+| [PFUAggDatabase](https://energyeconomydecoupling.github.io/PFUAggDatabase/) | The [targets](https://books.ropensci.org/targets/) pipeline to aggregate PSUT matrices created by the [PFUDatabase](https://energyeconomydecoupling.github.io/PFUDatabase/) pipeline |
 

### 2023-07-23 paper.Rmd
Working on the paper.

@@
 
-Historically, SEA has analyzed the ECCs of individual countries
+Historically, SEA has analyzed the ECCs of individual countries, 
+starting with primary and final stage data from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
+world energy balances,
 using a large number of linked spreadsheets, 
@@
 The PSUT framework succinctly describes the flow of energy carriers
-(``products'' in PSUT terminology) 
+("products" in PSUT terminology) 
 among energy conversion machines 
-(``industries'' in PSUT terminology)
-in a set of four matrices as described in the following table.
+("industries" in PSUT terminology)
+in a set of six matrices as described in the following table.
 
@@
 | **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy  |
+| **U_feed**  | product x industry | Feedstock use matrix           | Describes feedstock inputs to energy conversion devices  |
 | **U_EIOU**  | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
@@
 |:--------|:---------|
-| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulate row and column names in [matsindf](https://matthewheun.github.io/matsindf/) data frames |
-| [matsbyname](https://matthewheun.github.io/matsbyname/)  | Perform matrix mathematics that respects row and column names |
-| [matsindf](https://matthewheun.github.io/matsindf/)      | Store matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
+| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulates row and column names in [matsindf](https://matthewheun.github.io/matsindf/) data frames |
+| [matsbyname](https://matthewheun.github.io/matsbyname/)  | Performs matrix mathematics that respects row and column names |
+| [matsindf](https://matthewheun.github.io/matsindf/)      | Stores matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
 

### 2023-07-24 paper.Rmd
Better targets URL.

@@
 | [Recca](https://matthewheun.github.io/Recca/)       | Performs `R` energy conversion chain analysis |
-| [PFUDatabase](https://energyeconomydecoupling.github.io/PFUDatabase/) | The [targets](https://books.ropensci.org/targets/) pipeline to create a data frame of PSUT matrices |
-| [PFUAggDatabase](https://energyeconomydecoupling.github.io/PFUAggDatabase/) | The [targets](https://books.ropensci.org/targets/) pipeline to aggregate PSUT matrices created by the [PFUDatabase](https://energyeconomydecoupling.github.io/PFUDatabase/) pipeline |
+| [PFUDatabase](https://energyeconomydecoupling.github.io/PFUDatabase/) | A [targets](https://docs.ropensci.org/targets/) pipeline to create a data frame of PSUT matrices |
+| [PFUAggDatabase](https://energyeconomydecoupling.github.io/PFUAggDatabase/) | A [targets](https://docs.ropensci.org/targets/) pipeline to aggregate PSUT matrices |
 

### 2023-07-24 paper.Rmd
Added information about more packages.

@@
 
-Several decisions influenced the design and APIs for the packages.
-The decision involves data format. 
+The most important decision for the design of SEA packages
+involves data format.
 We selected the Physical Supply-Use Table (PSUT) framework,
 a matrix approach to describing energy flows from 
-extraction to
+resource extraction to
 processing stages and, ultimately, 
@@
 ("industries" in PSUT terminology)
-in a set of six matrices as described in the following table.
+with a set of six matrices as described in the following table.
 
@@
 | **R**       | industry x product | Resource matrix                | Contains exogeneous energy inputs to an ECC              |
-| **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy  |
+| **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy; the sum of **U_feed** and **U_EIOU** |
 | **U_feed**  | product x industry | Feedstock use matrix           | Describes feedstock inputs to energy conversion devices  |
@@
 
-The matrix approach to data storage brings the challenge that 
-different countries, 
-and different years for a given country,
+Further development followed from selection of the PSUT data format.
+First, the PSUT framework (a matrix-based data format) brought the challenge that 
+different countries and years 
 have varying energy energy carriers (products) and
 varying energy conversion machines (industries).
-To get around this challenge, a new `R` package 
-[matsbyname](https://matthewheun.github.io/matsbyname/)
-enables matrix mathematics that respects row and column names, 
+I.e., matrices for different countries and years have differing row and column names.
+To get around this challenge, we created an `R` package 
+([matsbyname](https://github.com/MatthewHeun/matsbyname/)) which
+enables matrix mathematics that respects matrix row and column names, 
 inserting `0` rows or columns where needed.
-
-Furthermore, 
-it would be convenient to perform matrix mathematics in `R` data frames
-as easily as scalar mathematics in data frames with [tidyverse](https://www.tidyverse.org) syntax. 
-[matsindf](https://matthewheun.github.io/matsindf/) enables this functionality.
+Second,
+we knew it would be convenient to perform _matrix_ mathematics 
+as easily as  _scalar_ mathematics
+in `R` data frames
+using the [tidyverse](https://www.tidyverse.org) syntax. 
+We developed the [matsindf](https://matthewheun.github.io/matsindf/) package 
+to enable this functionality.
+Finally, manipulating row and column names proved to be a challenge, 
+especially for matrices in PSUT data frames, so 
+we developed the [RCLabels](https://matthewheun.github.io/RCLabels/) package
+for that purpose.
+The table below summarizes these packages, 
+all of which are generally useful and available on 
+[CRAN](https://cran.r-project.org).
 
@@
 |:--------|:---------|
-| [RCLabels](https://matthewheun.github.io/RCLabels/)      | Manipulates row and column names in [matsindf](https://matthewheun.github.io/matsindf/) data frames |
-| [matsbyname](https://matthewheun.github.io/matsbyname/)  | Performs matrix mathematics that respects row and column names |
-| [matsindf](https://matthewheun.github.io/matsindf/)      | Stores matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
+| [RCLabels](https://github.com/MatthewHeun/RCLabels/)      | Manipulates row and column names in [matsindf](https://github.com/MatthewHeun/matsindf/) data frames |
+| [matsbyname](https://github.com/MatthewHeun/matsbyname/)  | Performs matrix mathematics that respects row and column names |
+| [matsindf](https://github.com/MatthewHeun/matsindf/)      | Stores matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
 
-The packages in the table above are applicable to many problem domains and
-are available on [CRAN](https://cran.r-project.org).
-
-Additional packages specific to ECC analysis within SEA are available on GitHub
-and described in the following table.
+Several calculation steps are required to create the PFU database. 
+Each step is assisted by an `R` package that we created.
+First, the IEA's primary- and final-stage WEEB data must be converted
+to the PSUT format, 
+a task completed by the [IEATools](https://github.com/MatthewHeun/IEATools/) package. 
+Second, human and animal muscle work must be calculated from 
+labor and FAO data following the methodology of @Steenwyk:2022ww
+using the [MWTools](https://github.com/EnergyEconomyDecoupling/MWTools) package. 
+Third, the IEA's primary- and final-stage ECC data 
+are extended to the useful stage by
+(a) allocating final stage energy to end-use machines and
+(b) multiplying allocated final energy by the 
+final-to-useful efficiency of each machine.
+This task is accomplished by the
+[PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/) 
+and 
+[Recca](https://github.com/MatthewHeun/Recca) 
+packages.
+Finally, ECCs must be converted from energy terms to exergy terms. 
+This step is assisted by the [Recca](https://github.com/MatthewHeun/Recca) package.
+The steps to create the PSUT matrices for each country and each year
+are accomplished by a
+[targets](https://docs.ropensci.org/targets/)
+computation pipeline
+available in the [PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/)
+package.
+These packages are specific to creating the PFU database 
+and are summarized in the table below.
 
@@
 |:--------|:---------|
-| [IEATools](https://matthewheun.github.io/IEATools/) | Converts IEA data to PSUT format |
+| [IEATools](https://github.com/MatthewHeun/IEATools/) | Converts IEA data to [matsindf](https://github.com/MatthewHeun/matsindf/) format |
+| [MWTools](https://github.com/energyeconomydecoupling/MWTools) | Converts ILO and FAO data to human and animal muscle work information in [matsindf](https://github.com/MatthewHeun/matsindf/) format |
 | [Recca](https://matthewheun.github.io/Recca/)       | Performs `R` energy conversion chain analysis |

### 2023-07-24 paper.Rmd
Editing.

@@
 Energy flows through societal energy conversion chains (ECCs)
-enable economic activity and facilitate human flourishing.
+to enable economic activity and facilitate human flourishing.
 To understand economic growth and human well-being, 
@@
 
-Historically, SEA has analyzed the ECCs of individual countries, 
+Historically, societal exergy analysts have
+analyzed the ECCs of individual countries
+using a large number of linked spreadsheets,
 starting with primary and final stage data from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
-world energy balances,
-using a large number of linked spreadsheets, 
-with data in varying and inconsistent formats 
+world energy balances
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
+Data were in varying and inconsistent formats.
 The authors of this paper and others in the field
@@
 deemed not scalable.
-We created a suite of `R` packages to provide scalable SEA 
-for every country in the world.
+We created a suite of `R` packages to create a database 
+of PFU data (the PFU Database)
+for all countries in the world from 1960--2020.
+The packages enable, for the first time, scalable SEA.
 This paper describes the design of those packages and
@@
 Further development followed from selection of the PSUT data format.
-First, the PSUT framework (a matrix-based data format) brought the challenge that 
+First, the matrix-based PSUT framework brought the challenge that 
 different countries and years 
 have varying energy energy carriers (products) and
-varying energy conversion machines (industries).
-I.e., matrices for different countries and years have differing row and column names.
-To get around this challenge, we created an `R` package 
-([matsbyname](https://github.com/MatthewHeun/matsbyname/)) which
-enables matrix mathematics that respects matrix row and column names, 
-inserting `0` rows or columns where needed.
+varying energy conversion machines (industries),
+meaning that PSUT matrices for different countries and years 
+have differing row and column names.
+To get around this challenge, we created the
+([matsbyname](https://github.com/MatthewHeun/matsbyname/)) 
+package which enables matrix mathematics 
+that respects matrix row and column names, 
+inserting rows or columns of `0`s when needed.
 Second,

### 2023-07-24 paper.Rmd
Working on paper

@@
 deemed not scalable.
-We created a suite of `R` packages to create a database 
+We created a suite of `R` packages to create a database^[Strictly speaking, we create data frames of matrices, 
+not a structured database.]
 of PFU data (the PFU Database)
@@
 meaning that PSUT matrices for different countries and years 
-have differing row and column names.
+have differing sizes and differing row and column names.
 To get around this challenge, we created the

### 2023-07-24 paper.Rmd
Completed section on PFUAggDatabase and related packages.

@@
 from the primary stage 
-(coal, oil, natural gas) 
+(resources extracted from the environment, 
+such as coal, oil, natural gas) 
 to the final stage
-(energy purchased by consumers, such as refined petroleum and electricity),
+(energy purchased by consumers, 
+such as refined petroleum and electricity),
 to the useful stage
-(heat, mechanical drive, and light),
+(energy desired by the end user, 
+such as heat, motion, and light),
 and sometimes to energy services
-(transport, thermal comfort, and illumination).
-Societal exergy analysis (SEA), an extension of energy analysis,
-quantifies ECCs as exergy, 
-the mechanical work potential of energy.
+(such as thermal comfort, transport, and illumination).
+Societal exergy analysis (SEA), 
+an extension of energy analysis,
+quantifies ECCs in exergy terms^[Exergy is 
+the mechanical work potential of energy.].
 
@@
 using a large number of linked spreadsheets,
-starting with primary and final stage data from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
+starting with primary- and final-stage data 
+from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
 world energy balances
@@
 wanted to expand SEA to cover all countries, but
-the spreadsheet approach to data analysis was
+the spreadsheet approach to SEA was
 deemed not scalable.
@@
 
-The most important decision for the design of SEA packages
+The most important decision for the suite of SEA packages
 involves data format.
-We selected the Physical Supply-Use Table (PSUT) framework,
+We authors are among those who
+developed the Physical Supply-Use Table (PSUT) framework,
 a matrix approach to describing energy flows from 
@@
 ("industries" in PSUT terminology)
-with a set of six matrices as described in the following table.
+with a set of six matrices described in the following table.
 
@@
 
-Further development followed from selection of the PSUT data format.
-First, the matrix-based PSUT framework brought the challenge that 
+Further development followed from selection of the PSUT framework.
+First, the matrices of the PSUT framework carry the challenge that 
 different countries and years 
-have varying energy energy carriers (products) and
+have varying energy carriers (products) and
 varying energy conversion machines (industries),
@@
 To get around this challenge, we created the
-([matsbyname](https://github.com/MatthewHeun/matsbyname/)) 
+[matsbyname](https://github.com/MatthewHeun/matsbyname/) [@Heun-matsbyname:2023]
 package which enables matrix mathematics 
@@
 in `R` data frames
-using the [tidyverse](https://www.tidyverse.org) syntax. 
-We developed the [matsindf](https://matthewheun.github.io/matsindf/) package 
+using the [tidyverse](https://www.tidyverse.org) syntax [@Wickham:2019]. 
+We developed the [matsindf](https://matthewheun.github.io/matsindf/) package [@Heun-matsindf:2023]
 to enable this functionality.
@@
 especially for matrices in PSUT data frames, so 
-we developed the [RCLabels](https://matthewheun.github.io/RCLabels/) package
+we developed the [RCLabels](https://matthewheun.github.io/RCLabels/) package [@Heun-RCLabels:2023]
 for that purpose.
@@
 to the PSUT format, 
-a task completed by the [IEATools](https://github.com/MatthewHeun/IEATools/) package. 
+a task completed by the [IEATools](https://github.com/MatthewHeun/IEATools/) package [@Heun-IEATools:2023]. 
 Second, human and animal muscle work must be calculated from 
 labor and FAO data following the methodology of @Steenwyk:2022ww
-using the [MWTools](https://github.com/EnergyEconomyDecoupling/MWTools) package. 
+using the [MWTools](https://github.com/EnergyEconomyDecoupling/MWTools) package [@Marshall:2023ab]. 
 Third, the IEA's primary- and final-stage ECC data 
@@
 This task is accomplished by the
-[PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/) 
+[Recca](https://github.com/MatthewHeun/Recca) [@Heun-Recca:2023]
 and 
-[Recca](https://github.com/MatthewHeun/Recca) 
+[PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/) [@Heun-PFUDatabase:2023]
 packages.
@@
 are accomplished by a
-[targets](https://docs.ropensci.org/targets/)
+[targets](https://docs.ropensci.org/targets/) [@Landau:2021aa]
 computation pipeline
@@
 package.
-These packages are specific to creating the PFU database 
-and are summarized in the table below.
+
+A second [targets](https://docs.ropensci.org/targets/) pipeline 
+in the [PFUAggDatabase](https://github.com/EnergyEconomyDecoupling/PFUAggDatabase) package [@Heun-PFUAggDatabase:2023]
+aggregates ECCs
+by region 
+(continents and world),
+by product category 
+(e.g., Coal and coal products, Oil and oil products, Low-temperature heat, etc.), 
+by end use categories
+(e.g., Residential, Transport, etc.), and
+to primary, final, and useful stages.
+Furthermore, 
+[PFUAggDatabase](https://github.com/EnergyEconomyDecoupling/PFUAggDatabase)
+calculates aggregated efficiencies.
+The [Recca](https://github.com/MatthewHeun/Recca) package is used extensively by the
+[PFUAggDatabase](https://github.com/EnergyEconomyDecoupling/PFUAggDatabase) 
+pipeline.
+
+The packages in the following table are specific to creating the PFU database 
+and are available on GitHub.
 

### 2023-07-24 paper.Rmd
More updates to the paper.

@@
 
+<!-- Define some macros to use later -->
+[link-cran]:           https://cran.r-project.org
+[link-ieatools]:       https://github.com/MatthewHeun/IEATools
+[link-ieaweeb]:        https://www.iea.org/data-and-statistics/data-product/world-energy-balances
+[link-matsbyname]:     https://github.com/MatthewHeun/matsbyname
+[link-matsindf]:       https://github.com/MatthewHeun/matsindf
+[link-mwtools]:        https://github.com/EnergyEconomyDecoupling/MWTools
+[link-pfudatabase]:    https://github.com/energyeconomydecoupling/PFUDatabase
+[link-pfuaggdatabase]: https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
+[link-rclabels]:       https://github.com/MatthewHeun/RCLabels
+[link-recca]:          https://github.com/MatthewHeun/Recca
+[link-targets]:        https://docs.ropensci.org/targets
+[link-tidyverse]:      https://www.tidyverse.org
+
 # Summary
 
-Energy flows through societal energy conversion chains (ECCs)
-to enable economic activity and facilitate human flourishing.
+Energy flowing through societal energy conversion chains (ECCs)
+enables economic activity and facilitates human flourishing.
 To understand economic growth and human well-being, 
-the field of energy analysis evaluates ECCs
+the field of societal energy analysis evaluates ECCs
 from the primary stage 
 (resources extracted from the environment, 
-such as coal, oil, natural gas) 
+such as coal, oil, and natural gas) 
 to the final stage
@@
 
+We created a suite of `R` packages 
+(some of which are generally useful and available through 
+[CRAN][link-cran],
+others of which are focused on SEA and available on GitHub)
+to assist SEA practitioners as they analyze 
+energy movement through society. 
+These new packages enable analysis of every country in the world
+across timespans of decades or longer.
+The new packages were used to create the "PFU Database," 
+a new resource for the SEA community. 
+The packages can be used for specific SEA analyses, too.
+
 # Statement of need
@@
 using a large number of linked spreadsheets,
-starting with primary- and final-stage data 
-from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
+often starting with primary- and final-stage data 
+from the [IEA][link-ieaweeb]'s
 world energy balances
@@
 To get around this challenge, we created the
-[matsbyname](https://github.com/MatthewHeun/matsbyname/) [@Heun-matsbyname:2023]
+[matsbyname][link-matsbyname] [@Heun-matsbyname:2023]
 package which enables matrix mathematics 
@@
 in `R` data frames
-using the [tidyverse](https://www.tidyverse.org) syntax [@Wickham:2019]. 
-We developed the [matsindf](https://matthewheun.github.io/matsindf/) package [@Heun-matsindf:2023]
+using the [tidyverse][link-tidyverse] syntax [@Wickham:2019]. 
+We developed the [matsindf][link-matsindf] package [@Heun-matsindf:2023]
 to enable this functionality.
@@
 especially for matrices in PSUT data frames, so 
-we developed the [RCLabels](https://matthewheun.github.io/RCLabels/) package [@Heun-RCLabels:2023]
+we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
 for that purpose.
@@
 all of which are generally useful and available on 
-[CRAN](https://cran.r-project.org).
+[CRAN][link-cran].
 
@@
 |:--------|:---------|
-| [RCLabels](https://github.com/MatthewHeun/RCLabels/)      | Manipulates row and column names in [matsindf](https://github.com/MatthewHeun/matsindf/) data frames |
-| [matsbyname](https://github.com/MatthewHeun/matsbyname/)  | Performs matrix mathematics that respects row and column names |
-| [matsindf](https://github.com/MatthewHeun/matsindf/)      | Stores matrices in cells of a data frame, thereby enabling [tidyverse](https://www.tidyverse.org) syntax |
+| [RCLabels][link-rclabels]     | Manipulates row and column names in [matsindf][link-matsindf] data frames |
+| [matsbyname][link-matsbyname] | Performs matrix mathematics that respects row and column names |
+| [matsindf][link-matsindf]     | Stores matrices in cells of a data frame, thereby enabling [tidyverse][link-tidyverse] syntax |
 
@@
 to the PSUT format, 
-a task completed by the [IEATools](https://github.com/MatthewHeun/IEATools/) package [@Heun-IEATools:2023]. 
+a task completed by the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
 Second, human and animal muscle work must be calculated from 
 labor and FAO data following the methodology of @Steenwyk:2022ww
-using the [MWTools](https://github.com/EnergyEconomyDecoupling/MWTools) package [@Marshall:2023ab]. 
+using the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
 Third, the IEA's primary- and final-stage ECC data 
@@
 This task is accomplished by the
-[Recca](https://github.com/MatthewHeun/Recca) [@Heun-Recca:2023]
+[Recca][link-recca] [@Heun-Recca:2023]
 and 
@@
 packages.
+The pipeline is modular, such that data for an country can be added at a later date
+to improve the database. 
 Finally, ECCs must be converted from energy terms to exergy terms. 
-This step is assisted by the [Recca](https://github.com/MatthewHeun/Recca) package.
+This step is assisted by the [Recca][link-recca] package.
 The steps to create the PSUT matrices for each country and each year
 are accomplished by a
-[targets](https://docs.ropensci.org/targets/) [@Landau:2021aa]
+[targets][link-targets] [@Landau:2021aa]
 computation pipeline
-available in the [PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/)
+available in the [PFUDatabase][link-pfudatabase]
 package.
+A unique feature of the
+[PFUDatabase][link-pfudatabase]
+pipeline is an exemplar system that allows
+analyses to proceed when allocation or efficiency data for a country are unavailable.
 
-A second [targets](https://docs.ropensci.org/targets/) pipeline 
-in the [PFUAggDatabase](https://github.com/EnergyEconomyDecoupling/PFUAggDatabase) package [@Heun-PFUAggDatabase:2023]
+A second [targets][link-targets] pipeline 
+in the [PFUAggDatabase][link-pfuaggdatabase] package [@Heun-PFUAggDatabase:2023]
 aggregates ECCs
@@
 by product category 
-(e.g., Coal and coal products, Oil and oil products, Low-temperature heat, etc.), 
-by end use categories
+(e.g., Coal and coal products, 
+Low-, Medium-, and High-temperature heat, etc.), 
+by end use sectors
 (e.g., Residential, Transport, etc.), and
-to primary, final, and useful stages.
+to primary, final, and useful (PFU) stages.
 Furthermore, 
-[PFUAggDatabase](https://github.com/EnergyEconomyDecoupling/PFUAggDatabase)
-calculates aggregated efficiencies.
-The [Recca](https://github.com/MatthewHeun/Recca) package is used extensively by the
-[PFUAggDatabase](https://github.com/EnergyEconomyDecoupling/PFUAggDatabase) 
+[PFUAggDatabase][link-pfuaggdatabase]
+calculates aggregated efficiencies across 
+the primary, final, and useful stages of all ECCs in the database.
+The [Recca][link-recca] package is used extensively by the
+[PFUAggDatabase][link-pfuaggdatabase]
 pipeline.
@@
 |:--------|:---------|
-| [IEATools](https://github.com/MatthewHeun/IEATools/) | Converts IEA data to [matsindf](https://github.com/MatthewHeun/matsindf/) format |
-| [MWTools](https://github.com/energyeconomydecoupling/MWTools) | Converts ILO and FAO data to human and animal muscle work information in [matsindf](https://github.com/MatthewHeun/matsindf/) format |
-| [Recca](https://matthewheun.github.io/Recca/)       | Performs `R` energy conversion chain analysis |
-| [PFUDatabase](https://energyeconomydecoupling.github.io/PFUDatabase/) | A [targets](https://docs.ropensci.org/targets/) pipeline to create a data frame of PSUT matrices |
-| [PFUAggDatabase](https://energyeconomydecoupling.github.io/PFUAggDatabase/) | A [targets](https://docs.ropensci.org/targets/) pipeline to aggregate PSUT matrices |
+| [IEATools][link-idatools]             | Converts IEA data to [matsindf][link-matsindf] format |
+| [MWTools][link-mwtools]               | Converts ILO and FAO data to human and animal muscle work information in [matsindf][link-matsindf] format |
+| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
+| [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |
+| [PFUAggDatabase][link-pfuaggdatabase] | A [targets][link-targets] pipeline to aggregate PSUT matrices |
 
-Reference for data repository [@Marshall:2023aa].
-
-`Gala` is an Astropy-affiliated Python package for galactic dynamics. Python
-enables wrapping low-level languages (e.g., C) for speed without losing
-flexibility or ease-of-use in the user-interface. The API for `Gala` was
-designed to provide a class-based and user-friendly interface to fast (C or
-Cython-optimized) implementations of common operations such as gravitational
-potential and force evaluation, orbit integration, dynamical transformations,
-and chaos indicators for nonlinear dynamics. `Gala` also relies heavily on and
-interfaces well with the implementations of physical units and astronomical
-coordinate systems in the `Astropy` package [@astropy] (`astropy.units` and
-`astropy.coordinates`).
-
-`Gala` was designed to be used by both astronomical researchers and by
-students in courses on gravitational dynamics or astronomy. It has already been
-used in a number of scientific publications [@Pearson:2017] and has also been
-used in graduate courses on Galactic dynamics to, e.g., provide interactive
-visualizations of textbook material [@Binney:2008]. The combination of speed,
-design, and support for Astropy functionality in `Gala` will enable exciting
-scientific explorations of forthcoming data releases from the *Gaia* mission
-[@gaia] by students and experts alike.
+Input data for the PFU Database can be found in @Marshall:2023aa.
+Access to the PFU Database can be obtained by correspondence 
+with author [PEB](mailto:P.E.Brockway@leeds.ac.uk) data^[Access to the PFU Database
+is restricted to those who have access to the [IEA WEEB data][link-ieaweeb].].
 
@@
 # Conclusion
+
+The new `R` packages
+
+enables
 

### 2023-07-24 paper.Rmd
Add DeSterck reference

@@
 energy movement through society. 
-These new packages enable analysis of every country in the world
+The new packages enable analysis of any country in the world
 across timespans of decades or longer.
-The new packages were used to create the "PFU Database," 
-a new resource for the SEA community. 
-The packages can be used for specific SEA analyses, too.
+In short, the packages enable, for the first time, scalable SEA.
+We used the new packages to create the "PFU Database," 
+a new resource for the SEA community^[Strictly speaking, 
+we create data frames of matrices, 
+not an SQL or similar database.].
+
+This paper describes the design of the new packages and
+demonstrates briefly their use.
 
@@
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
-Data were in varying and inconsistent formats.
+Data were stored in varying and inconsistent formats.
+A different attempt to create an SEA database
+estimated technical energy efficiencies 
+of end-use machines by economic status in a country,
+thereby precluding the use of the database for energy-economy studies [@De-Stercke:2014].
+
 The authors of this paper and others in the field
@@
 deemed not scalable.
-We created a suite of `R` packages to create a database^[Strictly speaking, we create data frames of matrices, 
-not a structured database.]
-of PFU data (the PFU Database)
-for all countries in the world from 1960--2020.
-The packages enable, for the first time, scalable SEA.
-This paper describes the design of those packages and
-demonstrates briefly their use.
+A new approach to SEA was needed, 
+one that scaled across all countries 
+for many years without relying on economic data
+to estimate technical efficiencies.
 

### 2023-07-24 paper.Rmd
"the current paper"

@@
 
-The authors of this paper and others in the field
+The authors of the current paper and others in the field
 wanted to expand SEA to cover all countries, but

### 2023-07-24 paper.Rmd
Working on paper.

@@
 To understand economic growth and human well-being, 
-the field of societal energy analysis evaluates ECCs
+the field of societal energy analysis (among other activities)
+evaluates ECCs
 from the primary stage 
@@
 In short, the packages enable, for the first time, scalable SEA.
-We used the new packages to create the "PFU Database," 
-a new resource for the SEA community^[Strictly speaking, 
+We used the new packages to create the "PFU Database,"^[Strictly speaking, 
 we create data frames of matrices, 
-not an SQL or similar database.].
+not an SQL or similar database.] 
+a new resource for the SEA community [@Marshall:2023aa].
 
@@
 
-Historically, societal exergy analysts have
+Historically, SEA practitioners have
 analyzed the ECCs of individual countries
-using a large number of linked spreadsheets,
+using linked spreadsheets,
 often starting with primary- and final-stage data 
@@
 Data were stored in varying and inconsistent formats.
-A different attempt to create an SEA database
+An early attempt to create a worldwide SEA database [@De-Stercke:2014]
 estimated technical energy efficiencies 
 of end-use machines by economic status in a country,
-thereby precluding the use of the database for energy-economy studies [@De-Stercke:2014].
+thereby precluding the use of the database for energy-economy studies.
 
@@
 to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
-The PSUT framework succinctly describes the flow of energy carriers
+We chose the PSUT framework as the data format for the `R` packages, 
+because it succinctly describes the flow of energy carriers
 ("products" in PSUT terminology) 
@@
 ("industries" in PSUT terminology)
-with a set of six matrices described in the following table.
+via a set of six matrices described in the following table.
 
@@
 To get around this challenge, we created the
-[matsbyname][link-matsbyname] [@Heun-matsbyname:2023]
-package which enables matrix mathematics 
+[matsbyname][link-matsbyname] package [@Heun-matsbyname:2023]
+which enables matrix mathematics 
 that respects matrix row and column names, 
@@
 Several calculation steps are required to create the PFU database. 
-Each step is assisted by an `R` package that we created.
+Each step is assisted by a new `R` package.
 First, the IEA's primary- and final-stage WEEB data must be converted
@@
 Second, human and animal muscle work must be calculated from 
-labor and FAO data following the methodology of @Steenwyk:2022ww
+[International Labor Organization](https://www.ilo.org/) (ILO) and
+[Food and Agriculture Organization](https://www.fao.org/) (FAO) data,
+following the methodology of @Steenwyk:2022ww, 
 using the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
@@
 packages.
-The pipeline is modular, such that data for an country can be added at a later date
-to improve the database. 
+The pipeline is modular, such that allocation and efficiency data
+for a country can be added at any time to improve the database. 
 Finally, ECCs must be converted from energy terms to exergy terms. 
@@
 (continents and world),
-by product category 
+by energy carrier (product) category 
 (e.g., Coal and coal products, 
 Low-, Medium-, and High-temperature heat, etc.), 
-by end use sectors
+by energy conversion machines (industries) and end use sectors
 (e.g., Residential, Transport, etc.), and
@@
 Furthermore, 
-[PFUAggDatabase][link-pfuaggdatabase]
-calculates aggregated efficiencies across 
+[PFUAggDatabase][link-pfuaggdatabase] functions
+calculate aggregated efficiencies across 
 the primary, final, and useful stages of all ECCs in the database.
@@
 
-The packages in the following table are specific to creating the PFU database 
+The packages in the following table are specific to SEA analyses
 and are available on GitHub.
@@
 Access to the PFU Database can be obtained by correspondence 
-with author [PEB](mailto:P.E.Brockway@leeds.ac.uk) data^[Access to the PFU Database
-is restricted to those who have access to the [IEA WEEB data][link-ieaweeb].].
+with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Access to the PFU Database
+is restricted to those who have access to [IEA WEEB data][link-ieaweeb].].
 

### 2023-07-24 paper.Rmd
Adding FAO and ILO links.

@@
 [link-ieaweeb]:        https://www.iea.org/data-and-statistics/data-product/world-energy-balances
+[link-ilo]:            https://www.ilo.org/
+[link-fao]:            https://www.fao.org/
 [link-matsbyname]:     https://github.com/MatthewHeun/matsbyname
@@
 Second, human and animal muscle work must be calculated from 
-[International Labor Organization](https://www.ilo.org/) (ILO) and
-[Food and Agriculture Organization](https://www.fao.org/) (FAO) data,
+[International Labor Organization][link-ilo] (ILO) and
+[Food and Agriculture Organization][link-fao] (FAO) data,
 following the methodology of @Steenwyk:2022ww, 
@@
 |:--------|:---------|
-| [IEATools][link-idatools]             | Converts IEA data to [matsindf][link-matsindf] format |
-| [MWTools][link-mwtools]               | Converts ILO and FAO data to human and animal muscle work information in [matsindf][link-matsindf] format |
+| [IEATools][link-ieatools]             | Converts IEA data to [matsindf][link-matsindf] format |
+| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to human and animal muscle work in [matsindf][link-matsindf] format |
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |

### 2023-07-24 paper.Rmd
Tidying paper.

@@
 Input data for the PFU Database can be found in @Marshall:2023aa.
-Access to the PFU Database can be obtained by correspondence 
+Access to the PFU Database can be obtained via correspondence 
 with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Access to the PFU Database
@@
 
-# Citations
-
-Citations to entries in paper.bib should be in
-[rMarkdown](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)
-format.
-
-If you want to cite a software repository URL (e.g. something on GitHub without a preferred
-citation) then you can do it with the example BibTeX entry below for @fidgit.
-
-For a quick reference, the following citation commands can be used:
-- `@author:2001`  ->  "Author et al. (2001)"
-- `[@author:2001]` -> "(Author et al., 2001)"
-- `[@author1:2001; @author2:2001]` -> "(Author1 et al., 2001; Author2 et al., 2002)"
-
 # Figures
@@
 
-We acknowledge contributions from Brigitta Sipocz, Syrtis Major, and Semyeong
-Oh, and support from Kathryn Johnston during the genesis of this project.
+Do we have any acknowledgements?
 

### 2023-07-24 paper.Rmd
Better footnote.

@@
 Access to the PFU Database can be obtained via correspondence 
-with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Access to the PFU Database
-is restricted to those who have access to [IEA WEEB data][link-ieaweeb].].
+with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains
+primary and final energy [IEA WEEB][link-ieaweeb] data, 
+access to the PFU Database
+is restricted to those who have access to [IEA WEEB][link-ieaweeb] data.].
 

### 2023-07-25 paper.Rmd
Adding Aramendia reference.

@@
 
-The most important decision for the suite of SEA packages
+The most important decision for the design of the suite of SEA packages
 involves data format.
@@
 developed the Physical Supply-Use Table (PSUT) framework,
-a matrix approach to describing energy flows from 
-resource extraction to
-processing stages and, ultimately, 
-to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
+a matrix-based approach to describing energy flows 
+("products" in PSUT terminology) 
+from resource extraction to
+processing stages
+("industries" in PSUT terminology)
+and, ultimately, 
+to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv].
 We chose the PSUT framework as the data format for the `R` packages, 
-because it succinctly describes the flow of energy carriers
-("products" in PSUT terminology) 
-among energy conversion machines 
-("industries" in PSUT terminology)
+because it succinctly describes an entire ECC
 via a set of six matrices described in the following table.
@@
 which enables matrix mathematics 
-that respects matrix row and column names, 
+that respects row and column names, 
 inserting rows or columns of `0`s when needed.
@@
 we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
-for that purpose.
+to assist.
 The table below summarizes these packages, 
@@
 
-| Package | Function |
-|:--------|:---------|
+| Package | Purpose |
+|:--------|:--------|
 | [RCLabels][link-rclabels]     | Manipulates row and column names in [matsindf][link-matsindf] data frames |
@@
 
-Several calculation steps are required to create the PFU database. 
+Four calculation steps are required to create a PFU database. 
 Each step is assisted by a new `R` package.
@@
 and 
-[PFUDatabase](https://github.com/energyeconomydecoupling/PFUDatabase/) [@Heun-PFUDatabase:2023]
+[PFUDatabase][link-pfudatabase] [@Heun-PFUDatabase:2023]
 packages.
-The pipeline is modular, such that allocation and efficiency data
+
+The [PFUDatabase][link-pfudatabase] pipeline is modular, 
+such that allocation and efficiency data
 for a country can be added at any time to improve the database. 
@@
 
-| Package | Function |
-|:--------|:---------|
+| Package | Purpose |
+|:--------|:--------|
 | [IEATools][link-ieatools]             | Converts IEA data to [matsindf][link-matsindf] format |

### 2023-07-25 paper.Rmd
Better refs for CRAN packages.

@@
 [link-fao]:            https://www.fao.org/
-[link-matsbyname]:     https://github.com/MatthewHeun/matsbyname
-[link-matsindf]:       https://github.com/MatthewHeun/matsindf
+[link-matsbyname]:     https://CRAN.R-project.org/package=matsbyname
+[link-matsindf]:       https://CRAN.R-project.org/package=matsindf
 [link-mwtools]:        https://github.com/EnergyEconomyDecoupling/MWTools
@@
 [link-pfuaggdatabase]: https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
-[link-rclabels]:       https://github.com/MatthewHeun/RCLabels
+[link-rclabels]:       https://CRAN.R-project.org/package=RCLabels
 [link-recca]:          https://github.com/MatthewHeun/Recca

### 2023-07-25 paper.Rmd
Added PFUSetup reference.

--- 2023-07-25 paper.Rmd
@@
 [link-fao]:            https://www.fao.org/
-[link-matsbyname]:     https://github.com/MatthewHeun/matsbyname
-[link-matsindf]:       https://github.com/MatthewHeun/matsindf
+[link-matsbyname]:     https://CRAN.R-project.org/package=matsbyname
+[link-matsindf]:       https://CRAN.R-project.org/package=matsindf
 [link-mwtools]:        https://github.com/EnergyEconomyDecoupling/MWTools
+[link-pfusetup]:       https://github.com/energyeconomydecoupling/PFUSetup
 [link-pfudatabase]:    https://github.com/energyeconomydecoupling/PFUDatabase
 [link-pfuaggdatabase]: https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
-[link-rclabels]:       https://github.com/MatthewHeun/RCLabels
+[link-rclabels]:       https://CRAN.R-project.org/package=RCLabels
 [link-recca]:          https://github.com/MatthewHeun/Recca
@@
 packages.
+Fourth, ECCs must be converted from energy terms to exergy terms. 
+This step is assisted by the [Recca][link-recca] package.
 
-The [PFUDatabase][link-pfudatabase] pipeline is modular, 
-such that allocation and efficiency data
-for a country can be added at any time to improve the database. 
-Finally, ECCs must be converted from energy terms to exergy terms. 
-This step is assisted by the [Recca][link-recca] package.
 The steps to create the PSUT matrices for each country and each year
@@
 analyses to proceed when allocation or efficiency data for a country are unavailable.
+The calculation pipeline in the [PFUDatabase][link-pfudatabase] packge 
+also allows allocation and efficiency data
+for any country to be added at any time to improve the database.
 
@@
 aggregates ECCs
-by region 
+(a) by region 
 (continents and world),
-by energy carrier (product) category 
+(b) by energy carrier (product) category 
 (e.g., Coal and coal products, 
 Low-, Medium-, and High-temperature heat, etc.), 
-by energy conversion machines (industries) and end use sectors
+(c) by energy conversion machines (industries) and end use sectors
 (e.g., Residential, Transport, etc.), and
-to primary, final, and useful (PFU) stages.
-Furthermore, 
-[PFUAggDatabase][link-pfuaggdatabase] functions
-calculate aggregated efficiencies across 
-the primary, final, and useful stages of all ECCs in the database.
-The [Recca][link-recca] package is used extensively by the
-[PFUAggDatabase][link-pfuaggdatabase]
-pipeline.
+(d) to primary, final, and useful (PFU) stages.
+The [PFUAggDatabase][link-pfuaggdatabase]
+pipeline uses the [Recca][link-recca] package extensively.
+Both [targets][link-targets] pipelines 
+(in the [PFUDatabase][link-pfudatabase] and
+[PFUAggDatabase][link-pfuaggdatabase] packages)
+benefit from the 
+[PFUSetup][link-pfusetup] package, which
+identifies a root directory and storage locations
+for input and output data.
 
@@
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
+| [PFUSetup][link-pfusetup]             | Setup for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
 | [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |

### 2023-07-25 paper.Rmd
Editing references.

@@
 benefit from the 
-[PFUSetup][link-pfusetup] package, which
+[PFUSetup][link-pfusetup] package [@Heun-pfusetup:2023], which
 identifies a root directory and storage locations

### 2023-07-25 paper.Rmd
Better description of PFUSetup.

--- 2023-07-25 paper.Rmd
@@
 benefit from the 
-[PFUSetup][link-pfusetup] package, which
-identifies a root directory and storage locations
-for input and output data.
+[PFUSetup][link-pfusetup] package [@Heun-pfusetup:2023], which
+identifies storage locations for input and output data.
 

### 2023-07-25 paper.Rmd
Better description of PFUSetup in the table.

@@
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
-| [PFUSetup][link-pfusetup]             | Setup for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
+| [PFUSetup][link-pfusetup]             | Establishes input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
 | [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |

### 2023-07-25 paper.Rmd
Same as last commit.

--- 2023-07-25 paper.Rmd
@@
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
-| [PFUSetup][link-pfusetup]             | Setup for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
+| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
 | [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |

### 2023-07-25 paper.Rmd
SEA --> R
Minor edit.

@@
 In short, the packages enable, for the first time, scalable SEA.
-We used the new packages to create the "PFU Database,"^[Strictly speaking, 
+We used the new packages to create the "PFU Database,"^[Technically speaking, 
 we create data frames of matrices, 
@@
 
-The most important decision for the design of the suite of SEA packages
+The most important decision for the design of the suite of `R` packages
 involves data format.

### 2023-07-25 paper.Rmd
Better footnote.

@@
 We used the new packages to create the "PFU Database,"^[Technically speaking, 
-we create data frames of matrices, 
+we create [matsindf][link-matsindf]-style data frames, 
 not an SQL or similar database.] 

### 2023-07-27 paper.Rmd
Added ScienceDirect reference.

@@
 quantifies ECCs in exergy terms^[Exergy is 
-the mechanical work potential of energy.].
+the mechanical work potential of energy.
+See any number of web references for an exergy primer,
+including @SciDirect:2023].
 

### 2023-07-28 paper.Rmd
Edits to paper.

--- 2023-07-25 paper.Rmd
@@
 (resources extracted from the environment, 
-such as coal, oil, and natural gas) 
+such as coal, oil, natural gas, wind, and solar) 
 to the final stage
@@
 (such as thermal comfort, transport, and illumination).
-Societal exergy analysis (SEA), 
-an extension of energy analysis,
+Societal _exergy_ analysis (SEA), 
+an extension of societal _energy_ analysis,
 quantifies ECCs in exergy terms^[Exergy is 
-the mechanical work potential of energy.].
+the mechanical work potential of energy.
+See any number of web references for an exergy primer,
+including @SciDirect:2023. 
+**** Do we have a better primer? ****].
 
 We created a suite of `R` packages 
-(some of which are generally useful and available through 
-[CRAN][link-cran],
-others of which are focused on SEA and available on GitHub)
-to assist SEA practitioners as they analyze 
+to assist SEA practitioners to analyze 
 energy movement through society. 
@@
 In short, the packages enable, for the first time, scalable SEA.
-We used the new packages to create the "PFU Database,"^[Technically speaking, 
-we create data frames of matrices, 
-not an SQL or similar database.] 
+We used the new packages to create the MR-PFU database,^["MR-PFU" 
+is an initialism for "Multi-Region
+Primary, Final, Useful."
+The MR-PFU database is multi-regional in the sense that
+it contains many countries as well as 
+continental and world aggregations.
+Technically speaking, 
+we create [matsindf][link-matsindf] data frames, 
+not SQL or similar databases.] 
 a new resource for the SEA community [@Marshall:2023aa].
@@
 Data were stored in varying and inconsistent formats.
-An early attempt to create a worldwide SEA database [@De-Stercke:2014]
-estimated technical energy efficiencies 
-of end-use machines by economic status in a country,
-thereby precluding the use of the database for energy-economy studies.
+An early SEA database [@De-Stercke:2014]
+estimated energy efficiencies 
+of end-use machines by the economic status of countries,
+thereby precluding its use for energy-economy studies.
 
@@
 for many years without relying on economic data
-to estimate technical efficiencies.
+to estimate machine efficiencies.
 
@@
 
-The most important decision for the design of the suite of `R` packages
-involves data format.
+The most important design decision for the suite of `R` packages
+involved data format.
 We authors are among those who
@@
 because it succinctly describes an entire ECC
-via a set of six matrices described in the following table.
+for one country and one year
+via a set of six matrices
+(the "RUVY" matrices).
+See the following table.
 
@@
 
-Further development followed from selection of the PSUT framework.
-First, the matrices of the PSUT framework carry the challenge that 
+Further development followed from selecting the PSUT framework.
+First, the RUVY matrices carry the challenge that 
 different countries and years 
@@
 varying energy conversion machines (industries),
-meaning that PSUT matrices for different countries and years 
-have differing sizes and differing row and column names.
+meaning that RUVY matrices 
+have differing row names, differing column names, and differing sizes.
 To get around this challenge, we created the
@@
 that respects row and column names, 
-inserting rows or columns of `0`s when needed.
+inserting **0** vectors when needed.
 Second,
-we knew it would be convenient to perform _matrix_ mathematics 
+wanted to be able to perform _matrix_ mathematics 
 as easily as _scalar_ mathematics
@@
 Finally, manipulating row and column names proved to be a challenge, 
-especially for matrices in PSUT data frames, so 
+especially for RUVY matrices in [matsindf][link-matsindf] data frames, so 
 we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
@@
 |:--------|:--------|
-| [RCLabels][link-rclabels]     | Manipulates row and column names in [matsindf][link-matsindf] data frames |
+| [RCLabels][link-rclabels]     | Manipulates matrix row and column names in [matsindf][link-matsindf] data frames |
 | [matsbyname][link-matsbyname] | Performs matrix mathematics that respects row and column names |
-| [matsindf][link-matsindf]     | Stores matrices in cells of a data frame, thereby enabling [tidyverse][link-tidyverse] syntax |
+| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analysis with [tidyverse][link-tidyverse] syntax |
 
-Four calculation steps are required to create a PFU database. 
-Each step is assisted by a new `R` package.
+Broadly speaking, 
+four calculation steps are required to create a PFU database. 
+Each step is completed by one or more new `R` packages.
 First, the IEA's primary- and final-stage WEEB data must be converted
-to the PSUT format, 
-a task completed by the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
+to RUVY matrices for each country and year, 
+a task accomplished by functions in the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
 Second, human and animal muscle work must be calculated from 
@@
 following the methodology of @Steenwyk:2022ww, 
-using the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
+using functions in the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
 Third, the IEA's primary- and final-stage ECC data 
-are extended to the useful stage by
+must be extended to the useful stage by
 (a) allocating final stage energy to end-use machines and
@@
 final-to-useful efficiency of each machine.
-This task is accomplished by the
+This task is accomplished by functions in the
 [Recca][link-recca] [@Heun-Recca:2023]
@@
 Fourth, ECCs must be converted from energy terms to exergy terms. 
-This step is assisted by the [Recca][link-recca] package.
+This step is performed by functions in the [Recca][link-recca] package.
 
@@
 The calculation pipeline in the [PFUDatabase][link-pfudatabase] packge 
-also allows allocation and efficiency data
-for any country to be added at any time to improve the database.
+allows allocation and efficiency data
+for any country and year to be added at any time to improve the database.
 
@@
 |:--------|:--------|
-| [IEATools][link-ieatools]             | Converts IEA data to [matsindf][link-matsindf] format |
-| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to human and animal muscle work in [matsindf][link-matsindf] format |
+| [IEATools][link-ieatools]             | Converts [IEA WEEB][link-ieaweeb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
+| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
 | [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
-| [PFUDatabase][link-pfudatabase]       | A [targets][link-targets] pipeline to create a data frame of PSUT matrices |
-| [PFUAggDatabase][link-pfuaggdatabase] | A [targets][link-targets] pipeline to aggregate PSUT matrices |
+| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
+| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
 
-Input data for the PFU Database can be found in @Marshall:2023aa.
-Access to the PFU Database can be obtained via correspondence 
+Input data for the PFU database can be found in @Marshall:2023aa.
+Access to the PFU database can be obtained via correspondence 
 with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains

### 2023-07-28 paper.Rmd
Add acknowledgements.

@@
 
+Create a MR-PFU database from the GHA and ZAF data contained in the IEATools package.
+Create the expectd folder structure that we use on the database.
+Store all necessary data in this repository.
+Execute the commands to produce the database, 
+albeit for only two countries and two years.
+
 # Conclusion
 
-The new `R` packages
+The new `R` packages enable the creation of databases of primary, final, and useful stage
+energy and exergy energy conversion chains.
+Such databases will enable advances in societal exergy analysis community
+and lead to greater insights on economic growth and human well-being.
 
-enables
+<!-- # Mathematics -->
+
+<!-- Single dollars ($) are required for inline mathematics e.g. $f(x) = e^{\pi/x}$ -->
+
+<!-- Double dollars make self-standing equations: -->
+
+<!-- $$\Theta(x) = \left\{\begin{array}{l} -->
+<!-- 0\textrm{ if } x < 0\cr -->
+<!-- 1\textrm{ else} -->
+<!-- \end{array}\right.$$ -->
+
+<!-- You can also use plain \LaTeX for equations -->
+<!-- \begin{equation}\label{eq:fourier} -->
+<!-- \hat{f}(\omega) = \int_{-\infty}^{\infty} f(x) e^{i\omega x} dx -->
+<!-- \end{equation} -->
+<!-- and refer to \autoref{eq:fourier} from text. -->
+
+<!-- # Figures -->
+
+<!-- Figures can be included like this: -->
+
+<!-- ![Caption for example figure.\label{fig:example}](figure.png) -->
+<!-- and referenced from text using \autoref{fig:example}. -->
+
+<!-- Figure sizes can be customized by adding an optional second parameter: -->
+
+<!-- ![Caption for example figure.](figure.png){ width=20% } -->
 
@@
 
-# Mathematics
-
-Single dollars ($) are required for inline mathematics e.g. $f(x) = e^{\pi/x}$
-
-Double dollars make self-standing equations:
-
-$$\Theta(x) = \left\{\begin{array}{l}
-0\textrm{ if } x < 0\cr
-1\textrm{ else}
-\end{array}\right.$$
-
-You can also use plain \LaTeX for equations
-\begin{equation}\label{eq:fourier}
-\hat{f}(\omega) = \int_{-\infty}^{\infty} f(x) e^{i\omega x} dx
-\end{equation}
-and refer to \autoref{eq:fourier} from text.
-
-# Figures
-
-Figures can be included like this:
-
-![Caption for example figure.\label{fig:example}](figure.png)
-and referenced from text using \autoref{fig:example}.
-
-Figure sizes can be customized by adding an optional second parameter:
-
-![Caption for example figure.](figure.png){ width=20% }
-
-# Acknowledgements
+MKH acknowledges the support of Calvin University for 
+a sabbatical during which he developed the PSUT framework
+and course releases during which he wrote many of the `R` packages
+described in this paper.
+ZM acknowledges ******. 
+EA acknowledges ******.
+PEB acknowledges ******.
 

### 2023-07-28 paper.Rmd
Better conclusion.

@@
 
-The new `R` packages enable the creation of databases of primary, final, and useful stage
-energy and exergy energy conversion chains.
-Such databases will enable advances in societal exergy analysis community
-and lead to greater insights on economic growth and human well-being.
+The new `R` packages described in this paper
+enable the creation of databases of primary, final, and useful
+energy and exergy conversion chains.
+Such databases will enable advances in the societal exergy analysis community
+and lead to greater insights 
+about the role of energy 
+in economic growth and human well-being.
 
@@
 
-Do we have any acknowledgements?
-
 # References

### 2023-07-28 paper.Rmd
Better example
More about length

@@
 
-Max length: 250--1000 words
+Max length: 250--1000 words. At the moment, we are at 1444 words.
+But I saw many JOSS papers are longer than 1000 words. 
+And our paper is describing many packages, whereas
+most JOSS papers describe only a single package.
+So I think we are going to be OK.
+I think we can own the issue in the cover letter, 
+saying that we are willing to consider splitting into
+two papers upon the editor's advice.
 
@@
 to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv].
-We chose the PSUT framework as the data format for the `R` packages, 
+We chose the PSUT framework as the data format for the new `R` packages, 
 because it succinctly describes an entire ECC
@@
 Broadly speaking, 
-four calculation steps are required to create a PFU database. 
-Each step is completed by one or more new `R` packages.
+four calculation steps are required to create a MR-PFU database. 
+Each step is assisted by functions in one or more new `R` packages.
 First, the IEA's primary- and final-stage WEEB data must be converted
@@
 Create a MR-PFU database from the GHA and ZAF data contained in the IEATools package.   
-Create the expectd folder structure that we use on the database.
+Create the folder structure for the MR-PFU database.   
 Store all necessary data in this repository.   

### 2023-07-28 paper.Rmd
More edits.
Added acknowledgements.

@@
 
-Further development followed from selecting the PSUT framework.
+Further development followed selection of the PSUT framework.
 First, the RUVY matrices carry the challenge that 
@@
 described in this paper.
-ZM acknowledges ******. 
-EA acknowledges ******.
-PEB acknowledges ******.
+ZM's and PEB's time was funded by 
+the UK Research Council under EPSRC Fellowship award EP/R024254/1.
+EA's time was funded by 
+the University of Leeds (School of Earth and Environment).
 

### 2023-07-28 paper.Rmd
New reference.

@@
 [link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
+[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
 [link-rclabels]:         https://CRAN.R-project.org/package=RCLabels
@@
 pipeline uses the [Recca][link-recca] package extensively.
+
 Both [targets][link-targets] pipelines 
@@
 benefit from the 
-[PFUSetup][link-pfusetup] package [@Heun-pfusetup:2023], which
-identifies storage locations for input and output data.
+[PFUSetup][link-pfusetup] [@Heun-PFUSetup:2023] and
+[PFUPipelineTools][link-pfupipelinetools] [@Heun-PFUPipelineTools:2023]
+packages. 
+[PFUSetup][link-pfusetup] identifies storage locations 
+for pipeline input and output data.
+[PFUPipelineTools][link-pfupipelinetools] provides functions and constants
+common to both pipelines.
 
@@
 | [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
+| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
 | [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
@@
 # References
+
+**** Change to Zenodo DOIs for all package references?. ****

### 2023-07-28 paper.Rmd
Now referring to GitHub instead of CRAN for some packages.

@@
 [link-fao]:              https://www.fao.org/
-[link-matsbyname]:       https://CRAN.R-project.org/package=matsbyname
-[link-matsindf]:         https://CRAN.R-project.org/package=matsindf
+[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
+[link-matsindf]:         https://github.com/MatthewHeun/matsindf
 [link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
@@
 [link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
-[link-rclabels]:         https://CRAN.R-project.org/package=RCLabels
+[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
 [link-recca]:            https://github.com/MatthewHeun/Recca

### 2023-07-28 paper.Rmd
Better footnote.

@@
 Primary, Final, Useful."
-The MR-PFU database is multi-regional in the sense that
+The MR-PFU database is "multi-regional" in the sense that
 it contains many countries as well as 

### 2023-07-28 paper.Rmd
Working on package URLs

@@
 [link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
-[link-pfusetup]:         https://github.com/energyeconomydecoupling/PFUSetup
-[link-pfudatabase]:      https://github.com/energyeconomydecoupling/PFUDatabase
+[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
+[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
 [link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase

### 2023-07-30 paper.Rmd
Working on example

@@
 ```{r}
-library(Recca)
+library(IEATools)
+library(PFUSetup)
+home_path <- "ExampleFolder"
+iea_folder_path <- file.path(home_path, "IEA extended energy balance data")
+iea_data_path <- file.path("IEA extended energy balance data", "GH-ZA-TJ-Extended-Energy-Balances-sample-2022.csv")
+setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 
+                                 dropbox_path = "",
+                                 project_path = "",
+                                 iea_folder_path = iea_folder_path,
+                                 iea_data_path = iea_data_path,
+                                 version = "v_example")
 ```

### 2023-07-30 paper.Rmd
Minor word change.

@@
 primary and final energy [IEA WEEB][link-ieaweeb] data, 
-access to the PFU Database
+use of the PFU Database
 is restricted to those who have access to [IEA WEEB][link-ieaweeb] data.].

### 2023-07-30 paper.Rmd
Adding README.txt file for IEA data.

@@
 ```{r}
-library(IEATools)
+# Available on CRAN
+# library(matsbyname)
+# library(matsindf)
+# library(RCLabels)
+
+# Available from MatthewHeun on GitHub
+# with devtools::install_github("MatthewHeun/xxxxxx")
+# library(IEATools)
+
+# Available from EnergyEconomyDecoupling on GitHub
+# with devtools::install_github("EnergyEconomyDecoupling/xxxxxx")
+# library(PFUPipelineTools)
 library(PFUSetup)
+
+# Use the PFUSetup package to establish paths for the example
 home_path <- "ExampleFolder"
 iea_folder_path <- file.path(home_path, "IEA extended energy balance data")
-iea_data_path <- file.path("IEA extended energy balance data", "GH-ZA-TJ-Extended-Energy-Balances-sample-2022.csv")
+iea_data_path <- file.path("IEA extended energy balance data",
+                           "GH-ZA-TJ-Extended-Energy-Balances-sample-2022.csv")
 setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 

### 2023-07-30 ExampleFolder/_targets_pfudatabase.R
Adding a _targets.R file

@@ -1,0 +1,95 @@
+library(magrittr)
+library(targets)
+library(PFUDatabase)
+# targets::tar_make() to run the pipeline
+# targets::tar_make_future(workers = 8) to execute across multiple cores.
+# targets::tar_read(<<target_name>>) to view the results.
+# targets::tar_destroy() to start over with everything.
+# targets::tar_make(callr_function = NULL) to debug.
+
+
+# Input data version
+version <- "v_example"
+
+# Custom parameters
+years <- c(1971, 2000)      # The years to be analyzed
+
+countries <- c("BHA", "ZAF")
+
+
+# Additional exemplar countries are countries which aren't included in the workflow
+# as individual countries, but from which allocation or efficiency data may be 
+# obtained and assigned to countries in the workflow using the exemplar system.
+additional_exemplar_countries <- c("AFRI", # Africa 
+                                   "ASIA", # Asia
+                                   "EURP", # Europe 
+                                   "MIDE", # Middle East
+                                   "NAMR", # North America
+                                   "OCEN", # Oceania 
+                                   "SAMR", # South America 
+                                   "BUNK") # Bunkers
+
+# Which type of matrix objects should be created?
+# "matrix" is the built-in matrix object in R.
+# "Matrix" will provide sparse matrices.
+matrix_class <- "Matrix"
+
+# Should we specify non-energy flows?
+specify_non_energy_flows <- TRUE
+
+# Should we apply fixes to the IEA data?
+apply_fixes <- TRUE
+
+# Should we do a release of the results?
+release <- FALSE
+
+# End user-adjustable parameters.
+
+# WRLD should not be in both countries and additional_exemplar_countries
+if (("WRLD" %in% countries) & ("WRLD" %in% additional_exemplar_countries)) {
+  # Remove WRLD from additional_exemplar_countries
+  additional_exemplar_countries <- additional_exemplar_countries[!(additional_exemplar_countries == "WRLD")]
+}
+
+# WRLD should always be in countries or in additional_exemplar_countries.
+if (!("WRLD" %in% countries) & !("WRLD" %in% additional_exemplar_countries)) {
+  # Add WRLD to additional_exemplar_countries
+  additional_exemplar_countries <- c("WRLD", additional_exemplar_countries)
+}
+
+# Set up for multithreaded work on the local machine.
+future::plan(future.callr::callr)
+
+# Set options for all targets.
+targets::tar_option_set(
[…]

### 2023-08-08 paper.Rmd
Added subscripts.

@@
 | Matrix        | rows x columns     | Name                           | Description                                              |
-|:------------|:-------------------|:-------------------------------|:---------------------------------------------------------|
+|:--------------|:-------------------|:-------------------------------|:---------------------------------------------------------|
 | **R**         | industry x product | Resource matrix                | Contains exogeneous energy inputs to an ECC              |
-| **U**       | product x industry | Use matrix                     | Describes how each energy conversion device uses energy; the sum of **U_feed** and **U_EIOU** |
-| **U_feed**  | product x industry | Feedstock use matrix           | Describes feedstock inputs to energy conversion devices  |
-| **U_EIOU**  | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
+| **U**         | product x industry | Use matrix                     | Describes how each energy conversion device uses energy; the sum of **U**~feed~ and **U**~EIOU~ |
+| **U**~feed~   | product x industry | Feedstock use matrix           | Describes feedstock inputs to energy conversion devices  |
+| **U**~EOIU~   | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
 | **V**         | industry x product | Make matrix                    | Describes how each energy conversion device makes energy |

### 2023-08-08 paper.Rmd
Working on EA suggestions.

@@
 We authors are among those who
-developed the Physical Supply-Use Table (PSUT) framework,
-a matrix-based approach to describing energy flows 
+developed the Physical Supply-Use Table (PSUT) framework
+[@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv],
+an adaptation of economic Input-Output (IO) analysis [@Miller:2009]
+to describe energy flows through society.
+The Physical Supply-Use Table (PSUT) framework 
+is a matrix-based approach that tracks energy flows 
 ("products" in PSUT terminology) 
@@
 and, ultimately, 
-to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv].
+to final demand
+in a way that is compatible with the IEA's world extended energy balances 
+[@International-Energy-Agency:2023aa].
 We chose the PSUT framework as the data format for the new `R` packages, 
@@
 that respects row and column names, 
-inserting **0** vectors when needed.
+inserting **0** row or column vectors when needed.
 Second,
-wanted to be able to perform _matrix_ mathematics 
+we wanted to be able to perform _matrix_ mathematics 
 as easily as _scalar_ mathematics
@@
 [PFUDatabase][link-pfudatabase]
-pipeline is an exemplar system that allows
+pipeline is an innovative exemplar system that allows
 analyses to proceed when allocation or efficiency data for a country are unavailable.
-The calculation pipeline in the [PFUDatabase][link-pfudatabase] packge 
+When allocations or efficiencies are missing,
+a string of exemplar countries are queried for similar information.
+For example, exemplars for Belgium are
+France, Europe, and ultimately, the world, in that order.
+The calculation pipeline in the [PFUDatabase][link-pfudatabase] package 
 allows allocation and efficiency data

### 2023-08-08 paper.Rmd
Working on EA's comments.

@@
 ---
-title: 'MR-PFU: A suite of R packages for energy conversion chain analysis'
+title: 'CL-PFU: A suite of R packages for energy conversion chain analysis'
 tags:
@@
 In short, the packages enable, for the first time, scalable SEA.
-We used the new packages to create the MR-PFU database,^["MR-PFU" 
-is an initialism for "Multi-Region
+We used the new packages to create the CL-PFU database,^["CL-PFU" 
+is an initialism for "Country-Level
 Primary, Final, Useful."
-The MR-PFU database is "multi-regional" in the sense that
+The CL-PFU database is "multi-regional" in the sense that
 it contains many countries as well as 
@@
 Broadly speaking, 
-four calculation steps are required to create a MR-PFU database. 
+four calculation steps are required to create a CL-PFU database. 
 Each step is assisted by functions in one or more new `R` packages.
@@
 
-Create a MR-PFU database from the GHA and ZAF data contained in the IEATools package.   
-Create the folder structure for the MR-PFU database.   
+Create a CL-PFU database from the GHA and ZAF data contained in the IEATools package.   
+Create the folder structure for the CL-PFU database.   
 Store all necessary data in this repository.   
@@
 in economic growth and human well-being.
+An example of early work using the CL-PFU database
+is the calculation of useful stage Energy Return On Investment for fossil fuels [a], 
+for which the ECCTools [b] and EROITools [c] packages were developed, 
+expanding the software capabilities introduced here."
+
+***** Add Emmanuel references *****
+
+[a]: Aramendia E., Brockway P., Taylor P., Norman J., Heun M., Marshall Z. (2023), “Chapter 4: Estimation of fossil fuels useful stage energy return on investment and implications for
+renewable energy systems” in Exploring the useful energy implications of the global energy transition: a net energy perspective, PhD thesis, University of Leeds.
+[b]: Emmanuel Aramendia and Matthew K. Heun. ECCTools: Tools for modifying the Energy Conversion Chain. R package version 0.1.5. 2022. doi: 10.5281/zenodo.7692105.
+[c]: Emmanuel Aramendia and Matthew K. Heun. EROITools: Tools for calculating and aggregating Energy Return On Investment values. Package version: v0.1.1. 2023. doi: 10 . 5281 / zenodo .7696869.
 

### 2023-08-08 paper.Rmd
Updates to EA packages references.

@@
 [link-cran]:             https://cran.r-project.org
+[link-ecctools]:         https://github.com/earamendia/ECCTools
+[link-eroitools]:        https://github.com/earamendia/EROITools
 [link-ieatools]:         https://github.com/MatthewHeun/IEATools
@@
 An example of early work using the CL-PFU database
-is the calculation of useful stage Energy Return On Investment for fossil fuels [a], 
-for which the ECCTools [b] and EROITools [c] packages were developed, 
-expanding the software capabilities introduced here."
-
-***** Add Emmanuel references *****
-
-[a]: Aramendia E., Brockway P., Taylor P., Norman J., Heun M., Marshall Z. (2023), “Chapter 4: Estimation of fossil fuels useful stage energy return on investment and implications for
-renewable energy systems” in Exploring the useful energy implications of the global energy transition: a net energy perspective, PhD thesis, University of Leeds.
-[b]: Emmanuel Aramendia and Matthew K. Heun. ECCTools: Tools for modifying the Energy Conversion Chain. R package version 0.1.5. 2022. doi: 10.5281/zenodo.7692105.
-[c]: Emmanuel Aramendia and Matthew K. Heun. EROITools: Tools for calculating and aggregating Energy Return On Investment values. Package version: v0.1.1. 2023. doi: 10 . 5281 / zenodo .7696869.
+is the calculation of useful stage 
+Energy Return On Investment for fossil fuels [@Aramendia:2023aa], 
+for which the [ECCTools][link-ecctools] [@Aramendia:2022aa] and
+[EROITools][link-eroitools] [@Aramendia:2022ab] packages were developed, 
+expanding the software capabilities introduced here.^[Note
+that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages are
+not part of this paper.]
 

### 2023-08-08 paper.Rmd
Updates to the paper.

@@
 is an initialism for "Country-Level
-Primary, Final, Useful."
+Primary, Final, and Useful."
 The CL-PFU database is "multi-regional" in the sense that
@@
 estimated energy efficiencies 
-of end-use machines by the economic status of countries,
+of end-use machines using the economic status of countries,
 thereby precluding its use for energy-economy studies.
@@
 have differing row names, differing column names, and differing sizes.
-To get around this challenge, we created the
+To circumvent this challenge, we created the
 [matsbyname][link-matsbyname] package [@Heun-matsbyname:2023]
@@
 When allocations or efficiencies are missing,
-a string of exemplar countries are queried for similar information.
+a string of exemplar countries are queried for the required information.
 For example, exemplars for Belgium are
-France, Europe, and ultimately, the world, in that order.
-The calculation pipeline in the [PFUDatabase][link-pfudatabase] package 
+France, Europe, and ultimately, the World, in that order.
+The design of the calculation pipeline in the [PFUDatabase][link-pfudatabase] package 
 allows allocation and efficiency data
-for any country and year to be added at any time to improve the database.
+for any country or year to be added at a later date 
+to improve the database.
 
@@
 (a) by region 
-(continents and world),
+(e.g., continents and world),
 (b) by energy carrier (product) category 
@@
 
-The packages in the following table are specific to SEA analyses
-and are available on GitHub.
+The packages in the following table are available on GitHub.
 
@@
 expanding the software capabilities introduced here.^[Note
-that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages are
-not part of this paper.]
+that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages were
+not reviewed for this paper.]
 

### 2023-08-08 paper.Rmd
Ignore _targets

@@
 # Available on CRAN
-# library(matsbyname)
-# library(matsindf)
-# library(RCLabels)
+library(targets)
 
@@
                                  version = "v_example")
+
+# Run the PFUDatabase pipeline
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+
+# Run the PFUAggDatabase pipeline
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))
 ```

### 2023-08-08 paper.Rmd
Working on example

@@
 # Run the PFUDatabase pipeline
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
 
 # Run the PFUAggDatabase pipeline
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))
+# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))
 ```

### 2023-08-08 paper.Rmd
Remove comment about DOIs.
Working on example.

@@
 
-# Use the PFUSetup package to establish paths for the example
-home_path <- "ExampleFolder"
-iea_folder_path <- file.path(home_path, "IEA extended energy balance data")
-iea_data_path <- file.path("IEA extended energy balance data",
-                           "GH-ZA-TJ-Extended-Energy-Balances-sample-2022.csv")
-setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 
-                                 dropbox_path = "",
-                                 project_path = "",
-                                 iea_folder_path = iea_folder_path,
-                                 iea_data_path = iea_data_path,
-                                 version = "v_example")
-
 # Run the PFUDatabase pipeline
-# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
 
@@
 # References
-
-**** Change to Zenodo DOIs for all package references?. ****

### 2023-08-08 ExampleFolder/_targets_pfudatabase.R

@@ -1,20 +1,30 @@
 library(magrittr)
-library(targets)
 library(PFUDatabase)
+library(PFUSetup)
 # targets::tar_make() to run the pipeline
 # targets::tar_make_future(workers = 8) to execute across multiple cores.
 # targets::tar_read(<<target_name>>) to view the results.
 # targets::tar_destroy() to start over with everything.
 # targets::tar_make(callr_function = NULL) to debug.
 
-
-# Input data version
-version <- "v_example"
+# Use the PFUSetup package to establish paths for the example
+home_path <- "ExampleFolder"
+iea_folder_path <- file.path(home_path, "IEA extended energy balance data")
+iea_data_path <- file.path("IEA extended energy balance data",
+                           "GHA-ZAF-TJ-Extended-Energy-Balances-sample-2022.csv")
+# Create the pfu_setup object used in the targets scripts 
+# _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
+pfu_setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 
+                                     dropbox_path = "",
+                                     project_path = "",
+                                     iea_folder_path = iea_folder_path,
+                                     iea_data_path = iea_data_path,
+                                     version = "v_example")
 
 # Custom parameters
 years <- c(1971, 2000)      # The years to be analyzed
 
-countries <- c("BHA", "ZAF")
+countries <- c("GHA", "ZAF")
 
 
 # Additional exemplar countries are countries which aren't included in the workflow
@@ -74,22 +84,22 @@ PFUDatabase::get_pipeline(countries = countries,
                           apply_fixes = apply_fixes,
                           years = years,
                           how_far = "all_targets",
-                          iea_data_path = PFUSetup::get_abs_paths(version = version)[["iea_data_path"]],
-                          country_concordance_path = PFUSetup::get_abs_paths(version = version)[["country_concordance_path"]],
-                          mw_concordance_path = PFUSetup::get_abs_paths(version = version)[["mw_concordance_path"]],
-                          amw_analysis_data_path = PFUSetup::get_abs_paths(version = version)[["amw_analysis_data_path"]],
-                          hmw_analysis_data_path = PFUSetup::get_abs_paths(version = version)[["hmw_analysis_data_path"]],
-                          phi_constants_path = PFUSetup::get_abs_paths(version = version)[["phi_constants_path"]],
+                          iea_data_path = pfu_setup[["iea_data_path"]],
[…]

### 2023-08-08 ExampleFolder/_targets_pfudatabase.R
Working on example.

@@ -1,6 +1,9 @@
-library(magrittr)
-library(PFUDatabase)
+# Available on CRAN
+
+# Available from EnergyEconomyDecoupling on GitHub
+# with devtools::install_github("EnergyEconomyDecoupling/xxxxxx")
 library(PFUSetup)
+
 # targets::tar_make() to run the pipeline
 # targets::tar_make_future(workers = 8) to execute across multiple cores.
 # targets::tar_read(<<target_name>>) to view the results.

### 2023-08-08 ExampleFolder/_targets_pfudatabase.R
Add library(PFUDatabase)

@@ -2,6 +2,7 @@
 
 # Available from EnergyEconomyDecoupling on GitHub
 # with devtools::install_github("EnergyEconomyDecoupling/xxxxxx")
+library(PFUDatabase)
 library(PFUSetup)
 
 # targets::tar_make() to run the pipeline

### 2023-08-08 paper.Rmd
Comment the pipeline command

@@
 ```{r}
-# Available on CRAN
 library(targets)
-
-# Available from MatthewHeun on GitHub
-# with devtools::install_github("MatthewHeun/xxxxxx")
-# library(IEATools)
-
-# Available from EnergyEconomyDecoupling on GitHub
-# with devtools::install_github("EnergyEconomyDecoupling/xxxxxx")
-# library(PFUPipelineTools)
-library(PFUSetup)
-
 # Run the PFUDatabase pipeline
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
 

### 2023-08-09 ExampleFolder/_targets_pfudatabase.R
dropbox_path --> cloud_storage_path

@@ -19,7 +19,7 @@ iea_data_path <- file.path("IEA extended energy balance data",
 # Create the pfu_setup object used in the targets scripts 
 # _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
 pfu_setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 
-                                     dropbox_path = "",
+                                     cloud_storage_path = "",
                                      project_path = "",
                                      iea_folder_path = iea_folder_path,
                                      iea_data_path = iea_data_path,

### 2023-08-09 ExampleFolder/_targets_pfudatabase.R
Update the targets pipeline script.

@@ -5,22 +5,12 @@
 library(PFUDatabase)
 library(PFUSetup)
 
-# targets::tar_make() to run the pipeline
-# targets::tar_make_future(workers = 8) to execute across multiple cores.
-# targets::tar_read(<<target_name>>) to view the results.
-# targets::tar_destroy() to start over with everything.
-# targets::tar_make(callr_function = NULL) to debug.
-
-# Use the PFUSetup package to establish paths for the example
-home_path <- "ExampleFolder"
-iea_folder_path <- file.path(home_path, "IEA extended energy balance data")
-iea_data_path <- file.path("IEA extended energy balance data",
-                           "GHA-ZAF-TJ-Extended-Energy-Balances-sample-2022.csv")
-# Create the pfu_setup object used in the targets scripts 
-# _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
-pfu_setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 
+# Create the pfu_setup object for file paths.
+# Setting home_path and cloud_storage_path to "" (empty string)
+# gives relative paths.
+pfu_setup <- PFUSetup::get_abs_paths(home_path = "", 
                                      cloud_storage_path = "",
-                                     project_path = "",
+                                     project_path = "ExampleFolder",
                                      iea_folder_path = iea_folder_path,
                                      iea_data_path = iea_data_path,
                                      version = "v_example")

### 2023-08-09 ExampleFolder/_targets_pfudatabase.R
Fixing the script

@@ -8,12 +8,14 @@ library(PFUSetup)
 # Create the pfu_setup object for file paths.
 # Setting home_path and cloud_storage_path to "" (empty string)
 # gives relative paths.
-pfu_setup <- PFUSetup::get_abs_paths(home_path = "", 
-                                     cloud_storage_path = "",
-                                     project_path = "ExampleFolder",
-                                     iea_folder_path = iea_folder_path,
-                                     iea_data_path = iea_data_path,
-                                     version = "v_example")
+pfu_setup <- PFUSetup::get_abs_paths(
+  home_path = "", 
+  cloud_storage_path = "",
+  project_path = "ExampleFolder", 
+  iea_data_path = file.path("ExampleFolder",
+                            "IEA extended energy balance data",
+                            "GHA-ZAF-TJ-Extended-Energy-Balances-sample-2022.csv"),
+  version = "v_example")
 
 # Custom parameters
 years <- c(1971, 2000)      # The years to be analyzed

### 2023-08-10 paper.Rmd
Fixing the script
Minor changes.

@@
 pipeline is an innovative exemplar system that allows
-analyses to proceed when allocation or efficiency data for a country are unavailable.
+analyses to proceed when allocation or efficiency data are unavailable for a country.
 When allocations or efficiencies are missing,
-a string of exemplar countries are queried for the required information.
+a string of exemplar countries or regions
+are queried for the required information.
 For example, exemplars for Belgium are
@@
 ```{r}
+# Download FAO data for animal muscle work
+source(file.path("ExampleFolder", "DownloadScripts", "download_amw_data.R"))
 library(targets)
 # Run the PFUDatabase pipeline
-# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
 

### 2023-08-10 ExampleFolder/DownloadScripts/download_amw_data.R
Adding DownloadScripts folder

@@ -1,0 +1,17 @@
+# This script uses the R package FAOSTAT to download data for the number of live
+# animals in the form of a
+
+library(FAOSTAT)
+
+# Establish constants
+download_location <- pfu_setup[["amw_analysis_data_path"]]
+# live_animals_code <- MWTools::fao_codes$live_animals_code
+live_animals_code <- "QCL"
+download_filename <- ""
+
+# Download .zip file containing data for the number of live animals into a specified folder
+fao_amw_data <- FAOSTAT::get_faostat_bulk(code = live_animals_code, data_folder = download_location)
+
+# Save downloaded file
+# saveRDS(object = fao_amw_data,
+#         file = PFUSetup::get_abs_paths()$fao_data_path)

### 2023-08-10 ExampleFolder/_targets_pfudatabase.R
Edits to pfudatabase script

@@ -12,18 +12,12 @@ library(PFUSetup)
 # targets::tar_make(callr_function = NULL) to debug.
 
 # Use the PFUSetup package to establish paths for the example
-home_path <- "ExampleFolder"
-iea_folder_path <- file.path(home_path, "IEA extended energy balance data")
-iea_data_path <- file.path("IEA extended energy balance data",
-                           "GHA-ZAF-TJ-Extended-Energy-Balances-sample-2022.csv")
 # Create the pfu_setup object used in the targets scripts 
 # _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
-pfu_setup <- PFUSetup::get_abs_paths(home_path = file.path("ExampleFolder"), 
-                                     dropbox_path = "",
-                                     project_path = "",
-                                     iea_folder_path = iea_folder_path,
-                                     iea_data_path = iea_data_path,
-                                     version = "v_example")
+pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
+                                     project_path = "ExampleFolder",
+                                     version = "v_example", 
+                                     iea_year = "2022")
 
 # Custom parameters
 years <- c(1971, 2000)      # The years to be analyzed

### 2023-08-10 paper.Rmd
Working on example

@@
 ```{r}
-# Download FAO data for animal muscle work
+# (1) Purchase, download, and store IEA data in correct location.
+# (2) Download FAO and ILO data for muscle work calculations.
 source(file.path("ExampleFolder", "DownloadScripts", "download_amw_data.R"))
-library(targets)
-# Run the PFUDatabase pipeline
+# (3) Run the PFUDatabase pipeline
 targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
 
- # Run the PFUAggDatabase pipeline
+# (4) Run the PFUAggDatabase pipeline
 # targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))

### 2023-08-10 ExampleFolder/DownloadScripts/download_amw_data.R

@@ -1,17 +1,0 @@
-# This script uses the R package FAOSTAT to download data for the number of live
-# animals in the form of a
-
-library(FAOSTAT)
-
-# Establish constants
-download_location <- pfu_setup[["amw_analysis_data_path"]]
-# live_animals_code <- MWTools::fao_codes$live_animals_code
-live_animals_code <- "QCL"
-download_filename <- ""
-
-# Download .zip file containing data for the number of live animals into a specified folder
-fao_amw_data <- FAOSTAT::get_faostat_bulk(code = live_animals_code, data_folder = download_location)
-
-# Save downloaded file
-# saveRDS(object = fao_amw_data,
-#         file = PFUSetup::get_abs_paths()$fao_data_path)

### 2023-08-10 ExampleFolder/DownloadScripts/download_mw_data.R

@@ -1,0 +1,62 @@
+# # This script uses the R package FAOSTAT to download data for the number of live
+# # animals in the form of a
+# 
+# library(FAOSTAT)
+# 
+# # Establish constants
+# download_location <- pfu_setup[["amw_analysis_data_path"]]
+# # live_animals_code <- MWTools::fao_codes$live_animals_code
+# live_animals_code <- "QCL"
+# download_filename <- ""
+# 
+# # Download .zip file containing data for the number of live animals into a specified folder
+# fao_amw_data <- FAOSTAT::get_faostat_bulk(code = live_animals_code, data_folder = download_location)
+# 
+# # Save downloaded file
+# # saveRDS(object = fao_amw_data,
+# #         file = PFUSetup::get_abs_paths()$fao_data_path)
+# 
+# 
+
+
+
+
+
+# This script uses the R package {FAOSTAT} to download data 
+# for the number of live animals and
+# the R package {Rilostat} to download data 
+# for the number of employed persons and number of working hours
+
+library(FAOSTAT)
+library(Rilostat)
+
+# Establish the version folder in which to save the raw data
+pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
+                                     project_path = "ExampleFolder",
+                                     version = "v_example", 
+                                     iea_year = "2022")
+# Establish FAO constants
+download_location <- tempdir()
+live_animals_code <- "QCL"
+# Download .zip file containing data for the number of live animals into a specified folder
+fao_amw_data <- FAOSTAT::get_faostat_bulk(code = live_animals_code,
+                                          data_folder = download_location)
+# Save downloaded data
+saveRDS(object = fao_amw_data, file = pfu_setup$fao_data_path)
+
+
+# Establish ILO constants
+working_hours_code <- MWTools::ilo_codes$working_hours_code
+employment_code <- MWTools::ilo_codes$employment_code
+# Download ILO data
+ilo_working_hours_data <- Rilostat::get_ilostat(id = working_hours_code,
+                                                quiet = TRUE) |>
+  Rilostat::label_ilostat(code = c("ref_area"))
+ilo_employment_data <- Rilostat::get_ilostat(id = employment_code,
+                                             quiet = TRUE) |>
+  Rilostat::label_ilostat(code = c("ref_area"))
+# Save downloaded data
+saveRDS(object = ilo_employment_data,
[…]

### 2023-08-10 ExampleFolder/_targets_pfudatabase.R

@@ -15,9 +15,7 @@ pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
 
 # Custom parameters
 years <- c(1971, 2000)      # The years to be analyzed
-
 countries <- c("GHA", "ZAF")
-
 
 # Additional exemplar countries are countries which aren't included in the workflow
 # as individual countries, but from which allocation or efficiency data may be 
@@ -85,7 +83,8 @@ PFUDatabase::get_pipeline(countries = countries,
                           # Temperature data not required for V1, argument set to NULL.
                           ceda_data_folder = NULL,
                           fao_data_path = pfu_setup[["fao_data_path"]],
-                          ilo_data_path = pfu_setup[["ilo_data_path"]],
+                          ilo_employment_data_path = pfu_setup[["ilo_employment_data_path"]],
+                          ilo_working_hours_data_path = pfu_setup[["ilo_working_hours_data_path"]],
                           machine_data_path = pfu_setup[["machine_data_folder"]],
                           exemplar_table_path = pfu_setup[["exemplar_table_path"]],
                           fu_analysis_folder = pfu_setup[["fu_analysis_folder"]],

### 2023-08-10 paper.Rmd
Chagne the reporter.

@@
 # (3) Run the PFUDatabase pipeline
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
+                  reporter = "summary")
 

### 2023-08-10 paper.Rmd
Fix download script name

--- 2023-08-10 paper.Rmd
@@
 # (2) Download FAO and ILO data for muscle work calculations.
-source(file.path("ExampleFolder", "DownloadScripts", "download_amw_data.R"))
+source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
 # (3) Run the PFUDatabase pipeline
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
+                  reporter = "summary")
 

### 2023-08-10 ExampleFolder/DownloadScripts/download_mw_data.R

@@ -1,32 +1,7 @@
-# # This script uses the R package FAOSTAT to download data for the number of live
-# # animals in the form of a
-# 
-# library(FAOSTAT)
-# 
-# # Establish constants
-# download_location <- pfu_setup[["amw_analysis_data_path"]]
-# # live_animals_code <- MWTools::fao_codes$live_animals_code
-# live_animals_code <- "QCL"
-# download_filename <- ""
-# 
-# # Download .zip file containing data for the number of live animals into a specified folder
-# fao_amw_data <- FAOSTAT::get_faostat_bulk(code = live_animals_code, data_folder = download_location)
-# 
-# # Save downloaded file
-# # saveRDS(object = fao_amw_data,
-# #         file = PFUSetup::get_abs_paths()$fao_data_path)
-# 
-# 
-
-
-
-
-
 # This script uses the R package {FAOSTAT} to download data 
 # for the number of live animals and
 # the R package {Rilostat} to download data 
 # for the number of employed persons and number of working hours
-
 library(FAOSTAT)
 library(Rilostat)
 

### 2023-08-10 paper.Rmd
Updated example.

@@
 # targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))
+
+# Look at the PSUT data frame
+psut <- targets::tar_read(PSUT)
+head(psut)
+# Each column of the psut data frame contains matrices
+psut$R[[1]]
 ```

### 2023-08-11 ExampleFolder/_targets_pfuaggdatabase.R
Working on PFUAggDatabase pipeline

@@ -1,0 +1,57 @@
+library(targets)
+
+# Use the PFUSetup package to establish paths for the example
+# Create the pfu_setup object used in the targets scripts 
+# _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
+pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
+                                     project_path = "ExampleFolder",
+                                     version = "v_example", 
+                                     iea_year = "2022")
+
+
+
+# Custom parameters
+countries <- c("GHA", "ZAF")
+years <- c(1971, 2000)  
+
+# Tells whether to do the R and Y chops.
+do_chops <- FALSE
+
+# Set the release to be used for input.
+psut_release <- "20230618T131003Z-4c70f"    # v1.1 (with Matrix objects)
+
+# Should we release the results?
+release <- FALSE
+
+
+# End user-adjustable parameters.
+
+
+# Set up for multithreaded work on the local machine.
+future::plan(future.callr::callr)
+
+# Set options for all targets.
+targets::tar_option_set(
+  packages = "PFUAggDatabase",
+  # Indicate that storage and retrieval of subtargets
+  # should be done by the worker thread,
+  # not the main thread.
+  # These options set defaults for all targets.
+  # Individual targets can override.
+  storage = "worker",
+  retrieval = "worker"
+)
+
+# Pull in the pipeline
+PFUAggDatabase::get_pipeline(countries = countries,
+                             years = years,
+                             do_chops = do_chops,
+                             psut_release = psut_release,
+                             aggregation_maps_path = pfu_setup[["aggregation_mapping_path"]],
+                             pipeline_releases_folder = pfu_setup[["pipeline_releases_folder"]],
+                             pipeline_caches_folder = pfu_setup[["pipeline_caches_folder"]],
+                             release = release)
+
+
+
+

### 2023-08-11 ExampleFolder/_targets_pfudatabase.R

@@ -14,8 +14,8 @@ pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
                                      iea_year = "2022")
 
 # Custom parameters
-years <- c(1971, 2000)      # The years to be analyzed
 countries <- c("GHA", "ZAF")
+years <- c(1971, 2000)  
 
 # Additional exemplar countries are countries which aren't included in the workflow
 # as individual countries, but from which allocation or efficiency data may be 

### 2023-08-11 ExampleFolder/_targets_pfuaggdatabase.R
Add the psut_release

@@ -18,7 +18,7 @@ years <- c(1971, 2000)
 do_chops <- FALSE
 
 # Set the release to be used for input.
-psut_release <- "20230618T131003Z-4c70f"    # v1.1 (with Matrix objects)
+psut_release <- "20230811T122618Z-3b7b3"
 
 # Should we release the results?
 release <- FALSE

### 2023-08-17 paper.Rmd
Working on PFUAggDatabase pipeline
Finished both pipelines.

--- 2023-08-10 paper.Rmd
@@
 [link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
+[link-pins]:             https://pins.rstudio.com
 [link-rclabels]:         https://github.com/MatthewHeun/RCLabels
@@
 Primary, Final, and Useful."
-The CL-PFU database is "multi-regional" in the sense that
-it contains many countries as well as 
+The CL-PFU database contains many countries as well as 
 continental and world aggregations.
 Technically speaking, 
-we create [matsindf][link-matsindf] data frames, 
+we create [matsindf][link-matsindf] data frames
+stored as `[pins][link-pins]` on a pinboard, 
 not SQL or similar databases.] 
@@
 
-```{r}
+```{r, eval=FALSE}
 # (1) Purchase, download, and store IEA data in correct location.
-# (2) Download FAO and ILO data for muscle work calculations.
+# (2) Download FAO and ILO data for muscle work calculations
+# via the included script.
+# This download is needed only once and
+# the results are now stored in the repository. 
 source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
-# (3) Run the PFUDatabase pipeline
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
-                  reporter = "summary")
-
-# (4) Run the PFUAggDatabase pipeline
-# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"))
 ```
 
-Create a CL-PFU database from the GHA and ZAF data contained in the IEATools package.   
-Create the folder structure for the CL-PFU database.   
-Store all necessary data in this repository.   
-Execute the commands to produce the database, 
-albeit for only two countries and two years.
+```{r, eval=FALSE}
+# (3) Run the PFUDatabase pipeline
+targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+# (4) Run the PFUAggDatabase pipeline
+# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"), callr_function = NULL)
+targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
+                         workers = 4)
+```
+
+```{r}
+# (5) Review results using the saved pins
+# Establish the pinboard
+pinboard <- file.path("ExampleFolder", "OutputData", "PipelineReleases") |> 
+  pins::board_folder(versioned = TRUE)
+# Look at the psut data frame
+# created by the PFUDatabase pipeline
+psut <- pinboard |>
+  pins::pin_read("psut", version = "20230811T122618Z-3b7b3")
+head(psut)
+# Each column of the psut data frame contains matrices.
+# The data are stored as sparse matrices,
+# but easier to see as regular matrices.
+psut$R[[1]] |> 
+  as.matrix()
+
+# Look at the aggregate PFU efficiency data frame
+# created by the PFUAggDatabase pipeline
+agg_eta_pfu <- pinboard |> 
+  pins::pin_read("agg_eta_pfu", version = "")
+head(agg_eta_pfu)
+```
 

### 2023-08-17 ExampleFolder/_targets_pfuaggdatabase.R

@@ -18,10 +18,11 @@ years <- c(1971, 2000)
 do_chops <- FALSE
 
 # Set the release to be used for input.
+# This is a pre-made pins release available in the repository.
 psut_release <- "20230811T122618Z-3b7b3"
 
 # Should we release the results?
-release <- FALSE
+release <- TRUE
 
 
 # End user-adjustable parameters.

### 2023-08-17 ExampleFolder/_targets_pfudatabase.R

@@ -40,7 +40,8 @@ specify_non_energy_flows <- TRUE
 # Should we apply fixes to the IEA data?
 apply_fixes <- TRUE
 
-# Should we do a release of the results?
+# Should we do a release of the results to a pinboard?
+# A release has already been made and is available in the repository.
 release <- FALSE
 
 # End user-adjustable parameters.

### 2023-08-17 paper.Rmd
Reference for pins.

@@
 
-Max length: 250--1000 words. At the moment, we are at 1444 words.
-But I saw many JOSS papers are longer than 1000 words. 
-And our paper is describing many packages, whereas
-most JOSS papers describe only a single package.
-So I think we are going to be OK.
-I think we can own the issue in the cover letter, 
-saying that we are willing to consider splitting into
-two papers upon the editor's advice.
+<!-- Max length: 250--1000 words. At the moment, we are at 1444 words. -->
+<!-- But I saw many JOSS papers are longer than 1000 words.  -->
+<!-- And our paper is describing many packages, whereas -->
+<!-- most JOSS papers describe only a single package. -->
+<!-- So I think we are going to be OK. -->
+<!-- I think we can own the issue in the cover letter,  -->
+<!-- saying that we are willing to consider splitting into -->
+<!-- two papers upon the editor's advice.  -->
 
@@
 [PFUAggDatabase][link-pfuaggdatabase] packages)
-benefit from the 
+produce data frames readable by the [pins][link-pins] package [@Silga:2023].
+Both pipelines benefit from the 
 [PFUSetup][link-pfusetup] [@Heun-PFUSetup:2023] and
@@
 # (4) Run the PFUAggDatabase pipeline
-# targets::tar_make(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"), callr_function = NULL)
 targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
@@
 agg_eta_pfu <- pinboard |> 
-  pins::pin_read("agg_eta_pfu", version = "")
-head(agg_eta_pfu)
+  pins::pin_read("agg_eta_pfu", version = "20230817T180346Z-45741")
+agg_eta_pfu |> 
+  # Filter and delete columns 
+  # to present a subset of exergy results in a data frame
+  dplyr::filter(Energy.type == "X", IEAMW == "Both", 
+                Product.aggregation == "Specified", 
+                Industry.aggregation == "Specified", 
+                GrossNet == "Net",
+                Country %in% c("GHA", "ZAF")) |> 
+  dplyr::mutate(
+    Method = NULL, 
+    Energy.type = NULL,
+    Product.aggregation = NULL,
+    Industry.aggregation = NULL,
+    IEAMW = NULL,
+    Chopped.mat = NULL, 
+    Chopped.var = NULL, 
+    GrossNet = NULL
+  )
 ```

### 2023-08-17 paper.Rmd
Minor changes to paper.

@@
 head(psut)
-# Each column of the psut data frame contains matrices.
-# The data are stored as sparse matrices,
-# but easier to see as regular matrices.
-psut$R[[1]] |> 
-  as.matrix()
+# Each column of the psut data frame contains sparse matrices.
+colnames(psut$R[[1]])
+psut$R[[1]]
 

### 2023-08-17 paper.Rmd
Updates to example.

@@
 # (1) Purchase, download, and store IEA data in correct location.
+#     Sample data are provided at the correct location in the repository.
 # (2) Download FAO and ILO data for muscle work calculations
 #     via the included script.
-# This download is needed only once and
-# the results are now stored in the repository. 
+#     This download step is needed only once and
+#     the data are stored in the repository.
 source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
@@
 ```{r, eval=FALSE}
-# (3) Run the PFUDatabase pipeline
+# (3) Run the PFUDatabase pipeline to create RUVY matrices.
 targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
-# (4) Run the PFUAggDatabase pipeline
+# (4) Run the PFUAggDatabase pipeline to aggregate 
+#     energy and calculate efficiencies..
 targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
@@
 ```{r}
-# (5) Review results using the saved pins
-# Establish the pinboard
+# (5) Review results using saved pins, which are stored in the repository.
+#     Establish the pinboard.
 pinboard <- file.path("ExampleFolder", "OutputData", "PipelineReleases") |> 
@@
 #     Look at the psut data frame
-# created by the PFUDatabase pipeline
+#     created by the PFUDatabase pipeline.
 psut <- pinboard |>
-  pins::pin_read("psut", version = "20230811T122618Z-3b7b3")
+  pins::pin_read(name = "psut", version = "20230811T122618Z-3b7b3")
 head(psut)
-# Each column of the psut data frame contains sparse matrices.
+#     Each column of the psut data frame contains sparse matrices, 
+#     the columns of which are suppressed on display.
+#     Show the column names here.
 colnames(psut$R[[1]])
+#     Show an example resource matrix.
 psut$R[[1]]
 
-# Look at the aggregate PFU efficiency data frame
-# created by the PFUAggDatabase pipeline
+#     Review the aggregate PFU efficiency data frame
+#     created by the PFUAggDatabase pipeline.
 agg_eta_pfu <- pinboard |> 
-  pins::pin_read("agg_eta_pfu", version = "20230817T180346Z-45741")
+  pins::pin_read(name = "agg_eta_pfu", version = "20230817T180346Z-45741")
 agg_eta_pfu |> 
 #     Filter and delete columns 
-  # to present a subset of exergy results in a data frame
+#     to present a subset of exergy results in a data frame.
+#     The EX.p, EX.f, and EX.u columns contain aggregate exergy at
+#     primary, final, and useful stages, respectively, in TJ.
+#     The eta_pf, eta_fu, and eta_pu columns contain aggregate
+#     primary-to-final, final-to-useful, and 
+#     primary-to-useful efficiencies.
   dplyr::filter(Energy.type == "X", IEAMW == "Both", 

### 2023-08-17 paper.Rmd
Add a message to editors.

@@
 
-<!-- Max length: 250--1000 words. At the moment, we are at 1444 words. -->
-<!-- But I saw many JOSS papers are longer than 1000 words.  -->
-<!-- And our paper is describing many packages, whereas -->
-<!-- most JOSS papers describe only a single package. -->
-<!-- So I think we are going to be OK. -->
-<!-- I think we can own the issue in the cover letter,  -->
-<!-- saying that we are willing to consider splitting into -->
+<!-- Message to editors: -->
+
+<!-- This paper describes a suite of packages -->
+<!-- to assist SEA practitioners to analyze  -->
+<!-- energy movement through society.  -->
+<!-- The new packages enable analysis of any country in the world -->
+<!-- across timespans of decades or longer. -->
+<!-- In short, the packages enable, for the first time,  -->
+<!-- scalable Societal Exergy Analysis (SEA). -->
+
+<!-- In an email exchange in late July 2021,  -->
+<!-- editor Afron Smith recommended that we  -->
+<!-- create a separate repository for the paper,  -->
+<!-- mentioning all of the individual packages in the manuscript. -->
+<!-- Thus, the paper is slightly longer than 1000 words. -->
+<!-- That said, we are willing to consider splitting into -->
 <!-- two papers upon the editor's advice. -->
+
+<!-- No portion of this paper has been submitted or published -->
+<!-- in another peer-reviewed journal.  -->
+<!-- A paper with a very different focus  -->
+<!-- (efficiency results from the full CL-PFU database) -->
+<!-- is in preparation. -->
+<!-- Our desire is that the results paper will reference -->
+<!-- this JOSS paper. -->
 
@@
 See any number of web references for an exergy primer,
-including @SciDirect:2023. 
-**** Do we have a better primer? ****].
+including @SciDirect:2023.].
 
@@
 we create [matsindf][link-matsindf] data frames
-stored as `[pins][link-pins]` on a pinboard, 
+stored as [pins][link-pins] on a pinboard, 
 not SQL or similar databases.] 
@@
 expanding the software capabilities introduced here.^[Note
-that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages were
-not reviewed for this paper.]
-
-<!-- # Mathematics -->
-
-<!-- Single dollars ($) are required for inline mathematics e.g. $f(x) = e^{\pi/x}$ -->
-
-<!-- Double dollars make self-standing equations: -->
-
-<!-- $$\Theta(x) = \left\{\begin{array}{l} -->
-<!-- 0\textrm{ if } x < 0\cr -->
-<!-- 1\textrm{ else} -->
-<!-- \end{array}\right.$$ -->
-
-<!-- You can also use plain \LaTeX for equations -->
-<!-- \begin{equation}\label{eq:fourier} -->
-<!-- \hat{f}(\omega) = \int_{-\infty}^{\infty} f(x) e^{i\omega x} dx -->
-<!-- \end{equation} -->
-<!-- and refer to \autoref{eq:fourier} from text. -->
-
-<!-- # Figures -->
-
-<!-- Figures can be included like this: -->
-
-<!-- ![Caption for example figure.\label{fig:example}](figure.png) -->
-<!-- and referenced from text using \autoref{fig:example}. -->
-
-<!-- Figure sizes can be customized by adding an optional second parameter: -->
-
-<!-- ![Caption for example figure.](figure.png){ width=20% } -->
+that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages 
+are not the subject of this paper.]
 

### 2023-08-17 paper.Rmd
Minor changes.

@@
 <!-- This paper describes a suite of packages -->
-<!-- to assist SEA practitioners to analyze  -->
+<!-- to assist Societal Exergy Analysis practitioners to analyze  -->
 <!-- energy movement through society.  -->
@@
 <!-- across timespans of decades or longer. -->
-<!-- In short, the packages enable, for the first time,  -->
-<!-- scalable Societal Exergy Analysis (SEA). -->
+<!-- In short, the packages enable, for the first time, scalable SEA. -->
 
@@
 <!-- (efficiency results from the full CL-PFU database) -->
-<!-- is in preparation. -->
-<!-- Our desire is that the results paper will reference -->
-<!-- this JOSS paper. -->
+<!-- is in preparation for the IOP journal -->
+<!-- Environmental Research: Energy. -->
+<!-- It is our desire that the Environmental Research: Energy -->
+<!-- paper will reference this JOSS paper. -->
 
-<!-- Define some macros to use later -->
+<!-- ---Matthew Kuperus Heun, 17 August 2023 -->
+
+<!-- Define some macros for later use -->
 [link-cran]:             https://cran.r-project.org

### 2023-08-17 paper.Rmd
Eliminate date

@@
   index: 2
-date: "20 July 2023"
+# date: "17 Aug 2023"
 bibliography: paper.bib

### 2023-08-17 paper.Rmd
Add Erstevag reference.

@@
 (such as thermal comfort, transport, and illumination).
-Societal _exergy_ analysis (SEA), 
+Societal _exergy_ analysis (SEA, @Ertesvaag:2001), 
 an extension of societal _energy_ analysis,

### 2023-08-17 paper.Rmd
Fix a reference.

@@
 that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages 
-are not the subject of this paper.]
+are not the subject of this paper and were not peer reviewed.]
 

### 2023-08-17 paper.Rmd
Final edits.

@@
 (such as thermal comfort, transport, and illumination).
-Societal _exergy_ analysis (SEA, @Ertesvaag:2001), 
+Societal _exergy_ analysis (SEA) [@Ertesvaag:2001], 
 an extension of societal _energy_ analysis,
@@
 four calculation steps are required to create a CL-PFU database. 
-Each step is assisted by functions in one or more new `R` packages.
-First, the IEA's primary- and final-stage WEEB data must be converted
+Each step is assisted by functions in one or more of the new `R` packages.
+First, the IEA's primary- and final-stage 
+World Extended Energy Balance (WEEB) data must be converted
 to RUVY matrices for each country and year, 
@@
 [Recca][link-recca] [@Heun-Recca:2023]
-and 
-[PFUDatabase][link-pfudatabase] [@Heun-PFUDatabase:2023]
-packages.
+package.
 Fourth, ECCs must be converted from energy terms to exergy terms. 
-This step is performed by functions in the [Recca][link-recca] package.
+This step is also performed by functions in the [Recca][link-recca] package.
 
@@
 (d) to primary, final, and useful (PFU) stages.
+In addition, the [PFUAggDatabase][link-pfuaggdatabase] pipeline 
+calculates aggregate efficiencies for each available region.
 The [PFUAggDatabase][link-pfuaggdatabase]
@@
 # (1) Purchase, download, and store IEA data in correct location.
-#     Sample data are provided at the correct location in the repository.
+#     Sample data are provided at the correct location in the repository
+#     for this paper.
 # (2) Download FAO and ILO data for muscle work calculations
@@
 #     This download step is needed only once and
-#     the data are stored in the repository.
+#     the data are stored in the repository 
+#     for this paper.
 source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
@@
 # (3) Run the PFUDatabase pipeline to create RUVY matrices.
-targets::tar_make(script = file.path("ExampleFolder", "_targets_pfudatabase.R"))
+targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
+                         workers = 4)
 # (4) Run the PFUAggDatabase pipeline to aggregate 
-#     energy and calculate efficiencies..
+#     energy and calculate efficiencies.
 targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),

### 2023-08-17 paper.Rmd
Adjust comments in message to editors.
Another line for the message to editors.

@@
 
+<!-- Thank you for considering our paper. -->
+
 <!-- ---Matthew Kuperus Heun, 17 August 2023 -->
@@
 In short, the packages enable, for the first time, scalable SEA.
-We used the new packages to create the CL-PFU database,^["CL-PFU" 
+We used the new packages to create the CL-PFU database^["CL-PFU" 
 is an initialism for "Country-Level
@@
 stored as [pins][link-pins] on a pinboard, 
-not SQL or similar databases.] 
+not SQL or similar databases.],
 a new resource for the SEA community [@Marshall:2023aa].

### 2023-08-18 paper.Rmd
Leeds University --> University of Leeds

@@
   index: 1
-- name: School of Earth and Environment, Leeds University
+- name: School of Earth and Environment, University of Leeds
   index: 2

### 2023-08-18 paper.Rmd
Delete "(among other activities)"

@@
 To understand economic growth and human well-being, 
-the field of societal energy analysis (among other activities)
+the field of societal energy analysis
 evaluates ECCs

### 2023-08-20 paper.Rmd
Minor edits.

@@
 across timespans of decades or longer.
-In short, the packages enable, for the first time, scalable SEA.
+In short, the new packages enable, for the first time, scalable SEA.
 We used the new packages to create the CL-PFU database^["CL-PFU" 
@@
 to final demand
-in a way that is compatible with the IEA's world extended energy balances 
+in a way that is compatible with the IEA's world extended energy balances (WEEB)
 [@International-Energy-Agency:2023aa].
@@
 meaning that RUVY matrices 
-have differing row names, differing column names, and differing sizes.
+have differing row names, differing column names, and differing sizes
+from one country to another and for different years within the same country.
 To circumvent this challenge, we created the
@@
 | [matsbyname][link-matsbyname] | Performs matrix mathematics that respects row and column names |
-| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analysis with [tidyverse][link-tidyverse] syntax |
+| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analyses with the [tidyverse][link-tidyverse] syntax |
 
@@
 First, the IEA's primary- and final-stage 
-World Extended Energy Balance (WEEB) data must be converted
+WEEB data must be converted
 to RUVY matrices for each country and year, 
@@
 [PFUDatabase][link-pfudatabase]
-pipeline is an innovative exemplar system that allows
+pipeline enabled by the [Recca][link-recca] package 
+is an innovative exemplar system that allows
 analyses to proceed when allocation or efficiency data are unavailable for a country.
@@
 Input data for the PFU database can be found in @Marshall:2023aa.
-Access to the PFU database can be obtained via correspondence 
+Access to the full CL-PFU database can be obtained via correspondence 
 with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains
@@
 is restricted to those who have access to [IEA WEEB][link-ieaweeb] data.].
+Researchers with access to the [IEA WEEB][link-ieaweeb] data
+can create their own CL-PFU database by following the example below.
 
@@
 ```{r, eval=FALSE}
-# (1) Purchase, download, and store IEA data in correct location.
-#     Sample data are provided at the correct location in the repository
-#     for this paper.
+# (1) Purchase, download, and store IEA WEEB data in the correct location.
+#     Sample data for two countries (Ghana and South Africa) 
+#     and two years (1971 and 2000) are included (with permission)
+#     at the correct location in the repository for this paper.
 # (2) Download FAO and ILO data for muscle work calculations
@@
 #     This download step is needed only once and
-#     the data are stored in the repository 
+#     downloaded data are stored in the repository 
 #     for this paper.
@@
 ```{r}
-# (5) Review results using saved pins, which are stored in the repository.
+# (5) Review results using the pins package.
 #     Establish the pinboard.
@@
 colnames(psut$R[[1]])
-#     Show an example resource matrix.
+#     Show an example resource (R) matrix.
 psut$R[[1]]
@@
 [EROITools][link-eroitools] [@Aramendia:2022ab] packages were developed, 
-expanding the software capabilities introduced here.^[Note
+expanding the software capabilities introduced here^[Note
 that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages 
-are not the subject of this paper and were not peer reviewed.]
+are not the subject of this paper and were not peer reviewed.].
 

### 2023-08-20 paper.Rmd
Minor edits.

@@
 [link-eroitools]:        https://github.com/earamendia/EROITools
+[link-iea]:              https://www.iea.org
 [link-ieatools]:         https://github.com/MatthewHeun/IEATools
-[link-ieaweeb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
+[link-ieaweb]:           https://www.iea.org/data-and-statistics/data-product/world-energy-balances
 [link-ilo]:              https://www.ilo.org/
@@
 We created a suite of `R` packages 
-to assist SEA practitioners to analyze 
-energy movement through society. 
+to assist SEA practitioners to analyze ECCs. 
 The new packages enable analysis of any country in the world
@@
 often starting with primary- and final-stage data 
-from the [IEA][link-ieaweeb]'s
+from the [IEA][link-iea]'s
 world energy balances 
+([WEB][link-ieaweb]s)
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
@@
 of end-use machines using the economic status of countries,
-thereby precluding its use for energy-economy studies.
+thereby precluding use for energy-economy studies.
 
@@
 to final demand
-in a way that is compatible with the IEA's world extended energy balances (WEEB)
+in a way that is compatible with the [IEA][link-iea]'s
+[WEB][link-ieaweb] data
 [@International-Energy-Agency:2023aa].
@@
 First, the IEA's primary- and final-stage 
-WEEB data must be converted
+WEB data must be converted
 to RUVY matrices for each country and year, 
@@
 |:--------|:--------|
-| [IEATools][link-ieatools]             | Converts [IEA WEEB][link-ieaweeb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
+| [IEATools][link-ieatools]             | Converts [IEA WEB][link-ieaweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
 | [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
@@
 with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains
-primary and final energy [IEA WEEB][link-ieaweeb] data, 
+primary and final energy [IEA WEB][link-ieaweb] data, 
 use of the PFU Database
-is restricted to those who have access to [IEA WEEB][link-ieaweeb] data.].
-Researchers with access to the [IEA WEEB][link-ieaweeb] data
-can create their own CL-PFU database by following the example below.
+is restricted to those who have access to [IEA][link-iea] [WEB][link-ieaweb] data.].
+Researchers with access to the 
+[IEA][link-iea]'s [WEB][link-ieaweb] 
+data can create their own CL-PFU database 
+by following the example below.
 
@@
 ```{r, eval=FALSE}
-# (1) Purchase, download, and store IEA WEEB data in the correct location.
+# (1) Purchase, download, and store IEA WEB data in the correct location.
 #     Sample data for two countries (Ghana and South Africa) 

### 2023-08-20 paper.Rmd
Added a comma!

@@
 (resources extracted from the environment, 
-such as coal, oil, natural gas, wind, and solar) 
+such as coal, oil, natural gas, wind, and solar),
 to the final stage

### 2023-08-20 paper.Rmd
More edits.

@@
 often starting with primary- and final-stage data 
-from the [IEA][link-iea]'s
-world energy balances 
-([WEB][link-ieaweb]s)
+from the [International Energy Agency][link-iea]'s (IEA's)
+[world energy balances][link-ieaweb] (WEBs)
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
@@
 of end-use machines using the economic status of countries,
-thereby precluding use for energy-economy studies.
+thereby precluding use in energy-economy studies.
 

### 2023-08-20 paper.Rmd
Add reference for energy analysis

@@
 To understand economic growth and human well-being, 
-the field of societal energy analysis
+the field of societal energy analysis [@Common:1976wn]
 evaluates ECCs

### 2023-08-20 paper.Rmd
More minor edits.

@@
 In addition, the [PFUAggDatabase][link-pfuaggdatabase] pipeline 
-calculates aggregate efficiencies for each available region.
+calculates efficiencies for each country
+and all available aggregations.
 The [PFUAggDatabase][link-pfuaggdatabase]
@@
 data can create their own CL-PFU database 
-by following the example below.
+following the example below.
 
@@
 energy and exergy conversion chains.
-Such databases will enable advances in the societal exergy analysis community
-and lead to greater insights 
+Such databases will enable advances in societal exergy analysis
+and lead to new insights 
 about the role of energy 
@@
 is the calculation of useful stage 
-Energy Return On Investment for fossil fuels [@Aramendia:2023aa], 
+Energy Return On Investment (EROI) for fossil fuels [@Aramendia:2023aa], 
 for which the [ECCTools][link-ecctools] [@Aramendia:2022aa] and

### 2023-08-20 paper.Rmd
More small changes.

@@
 in `R` data frames
-using the [tidyverse][link-tidyverse] syntax [@Wickham:2019]. 
+using [tidyverse][link-tidyverse] syntax [@Wickham:2019]. 
 We developed the [matsindf][link-matsindf] package [@Heun-matsindf:2023]
@@
 calculates efficiencies for each country
-and all available aggregations.
+at all available aggregations.
 The [PFUAggDatabase][link-pfuaggdatabase]
@@
 is restricted to those who have access to [IEA][link-iea] [WEB][link-ieaweb] data.].
-Researchers with access to the 
+If desired, researchers with access to the 
 [IEA][link-iea]'s [WEB][link-ieaweb] 

### 2023-08-20 paper.Rmd
Delete "the"; add a comma.

@@
 | [matsbyname][link-matsbyname] | Performs matrix mathematics that respects row and column names |
-| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analyses with the [tidyverse][link-tidyverse] syntax |
+| [matsindf][link-matsindf]     | Stores matrices in cells of data frames, thereby enabling analyses with [tidyverse][link-tidyverse] syntax |
 
@@
 #     via the included script.
-#     This download step is needed only once and
+#     This download step is needed only once, and
 #     downloaded data are stored in the repository 

### 2023-08-20 paper.Rmd
web --> internet
Noted open-source nature of packages.

@@
 the mechanical work potential of energy.
-See any number of web references for an exergy primer,
+See any number of internet references for an exergy primer,
 including @SciDirect:2023.].
 
-We created a suite of `R` packages 
+We created a suite of open-source `R` packages 
 to assist SEA practitioners to analyze ECCs. 

### 2023-08-20 paper.Rmd
Highlight open source in the conclusion.

@@
 
-The new `R` packages described in this paper
+The new open source `R` packages described in this paper
 enable the creation of databases of primary, final, and useful

### 2023-08-20 paper.Rmd
Not scalable in the conclusion.

--- 2023-08-20 paper.Rmd
@@
 
-The new `R` packages described in this paper
-enable the creation of databases of primary, final, and useful
-energy and exergy conversion chains.
+The new open source `R` packages described in this paper
+enable, for the first time, 
+creation of databases of primary, final, and useful
+energy and exergy conversion chains
+in a scalable manner.
 Such databases will enable advances in societal exergy analysis

### 2023-08-20 paper.Rmd
Update note to editors.

@@
 
-<!-- This paper describes a suite of packages -->
+<!-- This paper describes a suite of open source R packages -->
 <!-- to assist Societal Exergy Analysis practitioners to analyze -->

### 2023-08-20 paper.Rmd
Add "(SEA)"

--- 2023-08-20 paper.Rmd
@@
 
-<!-- This paper describes a suite of packages -->
-<!-- to assist Societal Exergy Analysis practitioners to analyze -->
+<!-- This paper describes a suite of open source R packages -->
+<!-- to assist Societal Exergy Analysis (SEA) practitioners to analyze -->
 <!-- energy movement through society. -->

### 2023-08-20 paper.Rmd
Add more notes for submission.

--- 2023-08-20 paper.Rmd
@@
 
+<!-- Title: -->
+<!-- CL-PFU: A suite of R packages for energy conversion chain analysis -->
+
+<!-- Software's GIT repository: -->
+<!-- https://github.com/EnergyEconomyDecoupling/CL-PFU-JOSS-Paper -->
+
+<!-- Name of git branch containing the paper, not the path: -->
+<!-- <<BLANK>> -->
+
+<!-- Software version: -->
+<!-- This submission covers several packages, each with its own current version. -->
+
+<!-- Type of submission: -->
+<!-- New submission -->
+
+<!-- Main subject of the paper: -->
+<!-- Software for societal exergy analysis -->
+
 <!-- Message to editors: -->
-
-<!-- This paper describes a suite of packages -->
-<!-- to assist Societal Exergy Analysis practitioners to analyze -->
+<!-- This paper describes a suite of open source R packages -->
+<!-- to assist Societal Exergy Analysis (SEA) practitioners to analyze -->
 <!-- energy movement through society. -->

### 2023-09-18 paper.Rmd
and --> and/or

@@
 Low-, Medium-, and High-temperature heat, etc.), 
-(c) by energy conversion machines (industries) and end use sectors
+(c) by energy conversion machines (industries) and/or end use sectors
 (e.g., Residential, Transport, etc.), and

### 2023-11-04 paper.Rmd
Add a sentence noting that MW must be converted to RUVY
Update notation around IEA EWEB data.

@@
 [link-ieatools]:         https://github.com/MatthewHeun/IEATools
-[link-ieaweb]:           https://www.iea.org/data-and-statistics/data-product/world-energy-balances
+[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
 [link-ilo]:              https://www.ilo.org/
@@
 from the [International Energy Agency][link-iea]'s (IEA's)
-[world energy balances][link-ieaweb] (WEBs)
+[extended world energy balances][link-ieaeweb] (EWEBs)
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
@@
 in a way that is compatible with the [IEA][link-iea]'s
-[WEB][link-ieaweb] data
+[EWEB][link-ieaeweb] data
 [@International-Energy-Agency:2023aa].
@@
 First, the IEA's primary- and final-stage 
-WEB data must be converted
+EWEB data must be converted
 to RUVY matrices for each country and year, 
@@
 using functions in the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
+The muscle work data must also be converted to RUVY matrices.
 Third, the IEA's primary- and final-stage ECC data 
@@
 |:--------|:--------|
-| [IEATools][link-ieatools]             | Converts [IEA WEB][link-ieaweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
+| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
 | [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
@@
 with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains
-primary and final energy [IEA WEB][link-ieaweb] data, 
+primary and final energy [IEA EWEB][link-ieaeweb] data, 
 use of the PFU Database
-is restricted to those who have access to [IEA][link-iea] [WEB][link-ieaweb] data.].
+is restricted to those who have access to [IEA][link-iea] [EWEB][link-ieaeweb] data.].
 If desired, researchers with access to the 
-[IEA][link-iea]'s [WEB][link-ieaweb] 
+[IEA][link-iea]'s [EWEB][link-ieaeweb] 
 data can create their own CL-PFU database 
@@
 ```{r, eval=FALSE}
-# (1) Purchase, download, and store IEA WEB data in the correct location.
+# (1) Purchase, download, and store IEA EWEB data in the correct location.
 #     Sample data for two countries (Ghana and South Africa) 

### 2023-11-27 paper.Rmd
Split lines to avoid bleeding into margin. Addresses #7.

@@
 # (3) Run the PFUDatabase pipeline to create RUVY matrices.
-targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
-                         workers = 4)
+targets::tar_make_future(
+  script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
+  workers = 4
+)
 # (4) Run the PFUAggDatabase pipeline to aggregate 
 #     energy and calculate efficiencies.
-targets::tar_make_future(script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
-                         workers = 4)
+targets::tar_make_future(
+  script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
+  workers = 4
+)
 ```

### 2023-11-28 paper.Rmd
head(as.data.frame(psut)) to avoid nonprinting "i"

@@
   pins::pin_read(name = "psut", version = "20230811T122618Z-3b7b3")
-head(psut)
+head(as.data.frame(psut))
 #     Each column of the psut data frame contains sparse matrices, 

### 2023-11-28 paper.Rmd
Restrict to 2 lines of output.

--- 2023-11-27 paper.Rmd
@@
   pins::pin_read(name = "psut", version = "20230811T122618Z-3b7b3")
-head(psut)
+head(as.data.frame(psut), 2)
 #     Each column of the psut data frame contains sparse matrices, 

### 2023-12-02 paper.Rmd
Remove PEB from author list. Didn't contribute to s/w.

@@
   orcid: 0000-0002-7438-214X
-  affiliation: 1
+  affiliation: 1,2
 - name: Zeke Marshall
@@
   affiliation: 2
-- name: Paul E. Brockway
-  orcid: 0000-0001-6925-8040
-  affiliation: 2
+# - name: Paul E. Brockway
+#   orcid: 0000-0001-6925-8040
+#   affiliation: 2
 - name: Emmanuel Aramendia

### 2023-12-02 paper.Rmd
Change title

--- 2023-11-28 paper.Rmd
@@
 ---
-title: 'CL-PFU: A suite of R packages for energy conversion chain analysis'
+title: 'CLPFUDatabase: A suite of R packages for energy conversion chain analysis'
 tags:

### 2023-12-02 paper.Rmd
Fix typos in the paper.

@@
 Access to the full CL-PFU database can be obtained via correspondence 
-with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains
+with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the CL-PFU database contains
 primary and final energy [IEA EWEB][link-ieaeweb] data, 
-use of the PFU Database
+use of the CL-PFU database
 is restricted to those who have access to [IEA][link-iea] [EWEB][link-ieaeweb] data.].

### 2023-12-11 README.Rmd
Add Readme.md document.

--- 2023-12-02 paper.Rmd
@@
 ---
-title: 'CL-PFU: A suite of R packages for energy conversion chain analysis'
-tags:
-- energy
-- exergy
-- R
-- matrix mathematics
-- energy efficiency
-authors:
-- name: Matthew Kuperus Heun
-  orcid: 0000-0002-7438-214X
-  affiliation: 1,2
-- name: Zeke Marshall
-  orcid: 0000-0001-9260-7827
-  affiliation: 2
-- name: Paul E. Brockway
-  orcid: 0000-0001-6925-8040
-  affiliation: 2
-- name: Emmanuel Aramendia
-  orcid: 0000-0002-5964-6776
-  affiliation: 2
-affiliations:
-- name: Engineering Department, Calvin University
-  index: 1
-- name: School of Earth and Environment, University of Leeds
-  index: 2
-# date: "17 Aug 2023"
-bibliography: paper.bib
-output:
-  html_document:
-    keep_md: yes
+output: github_document
+bibliography: References.bib
 ---
 
-<!-- Title: -->
-<!-- CL-PFU: A suite of R packages for energy conversion chain analysis -->
-
-<!-- Software's GIT repository: -->
-<!-- https://github.com/EnergyEconomyDecoupling/CL-PFU-JOSS-Paper -->
-
-<!-- Name of git branch containing the paper, not the path: -->
-<!-- <<BLANK>> -->
-
-<!-- Software version: -->
-<!-- This submission covers several packages, each with its own current version. -->
-
-<!-- Type of submission: -->
-<!-- New submission -->
-
-<!-- Main subject of the paper: -->
-<!-- Software for societal exergy analysis -->
-
-<!-- Message to editors: -->
-<!-- This paper describes a suite of open source R packages -->
-<!-- to assist Societal Exergy Analysis (SEA) practitioners to analyze -->
-<!-- energy movement through society. -->
-<!-- The new packages enable analysis of any country in the world -->
-<!-- across timespans of decades or longer. -->
-<!-- In short, the packages enable, for the first time, scalable SEA. -->
-
-<!-- In an email exchange in late July 2021, -->
-<!-- editor Afron Smith recommended that we -->
-<!-- create a separate repository for the paper, -->
-<!-- mentioning all of the individual packages in the manuscript. -->
-<!-- Thus, the paper is slightly longer than 1000 words. -->
-<!-- That said, we are willing to consider splitting into -->
-<!-- two papers upon the editor's advice. -->
-
-<!-- No portion of this paper has been submitted or published -->
-<!-- in another peer-reviewed journal. -->
-<!-- A paper with a very different focus -->
-<!-- (efficiency results from the full CL-PFU database) -->
-<!-- is in preparation for the IOP journal -->
-<!-- Environmental Research: Energy. -->
-<!-- It is our desire that the Environmental Research: Energy -->
-<!-- paper will reference this JOSS paper. -->
-
-<!-- Thank you for considering our paper. -->
-
-<!-- ---Matthew Kuperus Heun, 17 August 2023 -->
+<!-- *********** -->
+<!-- Note: README.md is generated from README.Rmd.   -->
+<!-- Be sure to edit README.Rmd and generate the README.md file by Cmd/Ctl-shift-K -->
+<!-- *********** -->
 
@@
 
-# Summary
+## JOSS paper
 
-Energy flowing through societal energy conversion chains (ECCs)
-enables economic activity and facilitates human flourishing.
-To understand economic growth and human well-being, 
-the field of societal energy analysis [@Common:1976wn]
-evaluates ECCs
-from the primary stage 
-(resources extracted from the environment, 
-such as coal, oil, natural gas, wind, and solar),
-to the final stage
-(energy purchased by consumers, 
-such as refined petroleum and electricity),
-to the useful stage
-(energy desired by the end user, 
-such as heat, motion, and light),
-and sometimes to energy services
-(such as thermal comfort, transport, and illumination).
-Societal _exergy_ analysis (SEA) [@Ertesvaag:2001], 
-an extension of societal _energy_ analysis,
-quantifies ECCs in exergy terms^[Exergy is 
-the mechanical work potential of energy.
-See any number of internet references for an exergy primer,
-including @SciDirect:2023.].
-
-We created a suite of open-source `R` packages 
-to assist SEA practitioners to analyze ECCs. 
-The new packages enable analysis of any country in the world
-across timespans of decades or longer.
-In short, the new packages enable, for the first time, scalable SEA.
-We used the new packages to create the CL-PFU database^["CL-PFU" 
+This repository contains a paper submitted to 
+the [Journal of Open Source Software (JOSS)](https://joss.theoj.org)
+covering the suite of `R` packages associated with the 
+CL-PFU database^["CL-PFU" 
 is an initialism for "Country-Level
@@
 stored as [pins][link-pins] on a pinboard, 
-not SQL or similar databases.],
-a new resource for the SEA community [@Marshall:2023aa].
+not SQL or similar databases.].
 
-This paper describes the design of the new packages and
-demonstrates briefly their use.
+## CL-PFU database
 
-# Statement of need
+We created a suite of `R` packages 
+to support the creation and use of the CL-PFU database.
+The CL-PFU database is a new resouce 
+for the Societal Exergy Analysis (SEA) research community and
+assists SEA practitioners 
+to analyze energy conversion chains (ECCs).
+The new packages enable, for the first time, scalable SEA.
 
-Historically, SEA practitioners have
-analyzed the ECCs of individual countries
-using linked spreadsheets,
-often starting with primary- and final-stage data 
-from the [International Energy Agency][link-iea]'s (IEA's)
-[extended world energy balances][link-ieaeweb] (EWEBs)
-[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
-Data were stored in varying and inconsistent formats.
-An early SEA database [@De-Stercke:2014]
-estimated energy efficiencies 
-of end-use machines using the economic status of countries,
-thereby precluding use in energy-economy studies.
+## `R` packages
 
-The authors of the current paper and others in the field
-wanted to expand SEA to cover all countries, but
-the spreadsheet approach to SEA was
-deemed not scalable.
-A new approach to SEA was needed, 
-one that scaled across all countries 
-for many years without relying on economic data
-to estimate machine efficiencies.
-
-# Design of `R` packages
-
-The most important design decision for the suite of `R` packages
-involved data format.
-We authors are among those who
-developed the Physical Supply-Use Table (PSUT) framework
-[@Rocco:2016; @Guevara:2017; @Heun:2018; @Aramendia:2022tv],
-an adaptation of economic Input-Output (IO) analysis [@Miller:2009]
-to describe energy flows through society.
-The Physical Supply-Use Table (PSUT) framework 
-is a matrix-based approach that tracks energy flows 
-("products" in PSUT terminology) 
-from resource extraction to
-processing stages
-("industries" in PSUT terminology)
-and, ultimately, 
-to final demand
-in a way that is compatible with the [IEA][link-iea]'s
-[EWEB][link-ieaeweb] data
-[@International-Energy-Agency:2023aa].
-We chose the PSUT framework as the data format for the new `R` packages, 
-because it succinctly describes an entire ECC
-for one country and one year
-via a set of six matrices
-(the "RUVY" matrices).
-See the following table.
-
-| Matrix        | rows x columns     | Name                           | Description                                              |
-|:--------------|:-------------------|:-------------------------------|:---------------------------------------------------------|
-| **R**         | industry x product | Resource matrix                | Contains exogeneous energy inputs to an ECC              |
-| **U**         | product x industry | Use matrix                     | Describes how each energy conversion device uses energy; the sum of **U**~feed~ and **U**~EIOU~ |
-| **U**~feed~   | product x industry | Feedstock use matrix           | Describes feedstock inputs to energy conversion devices  |
-| **U**~EOIU~   | product x industry | Energy industry own use matrix | Describes how the energy industry uses energy            |
-| **V**         | industry x product | Make matrix                    | Describes how each energy conversion device makes energy |
-| **Y**         | product x industry | Final demand matrix            | Describes how each energy carrier is consumed            |
-
-Further development followed selection of the PSUT framework.
-First, the RUVY matrices carry the challenge that 
-different countries and years 
-have varying energy carriers (products) and
-varying energy conversion machines (industries),
-meaning that RUVY matrices 
-have differing row names, differing column names, and differing sizes
-from one country to another and for different years within the same country.
-To circumvent this challenge, we created the
-[matsbyname][link-matsbyname] package [@Heun-matsbyname:2023]
-which enables matrix mathematics 
-that respects row and column names, 
-inserting **0** row or column vectors when needed.
-Second,
-we wanted to be able to perform _matrix_ mathematics 
-as easily as _scalar_ mathematics
-in `R` data frames
-using [tidyverse][link-tidyverse] syntax [@Wickham:2019]. 
-We developed the [matsindf][link-matsindf] package [@Heun-matsindf:2023]
-to enable this functionality.
-Finally, manipulating row and column names proved to be a challenge, 
-especially for RUVY matrices in [matsindf][link-matsindf] data frames, so 
-we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
-to assist.
-The table below summarizes these packages, 
-all of which are generally useful and available on 
-[CRAN][link-cran].
+The `R` packages covered by this paper include
+three packages that support the PSUT framework of
+[Heun et al. [-@Heun:2018]](https://doi.org/10.1016/j.apenergy.2018.05.109).
+These packages are 
+generally useful and available on [CRAN][link-cran].
 
@@
 
-Broadly speaking, 
-four calculation steps are required to create a CL-PFU database. 
-Each step is assisted by functions in one or more of the new `R` packages.
-First, the IEA's primary- and final-stage 
-EWEB data must be converted
-to RUVY matrices for each country and year, 
-a task accomplished by functions in the [IEATools][link-ieatools] package [@Heun-IEATools:2023]. 
-Second, human and animal muscle work must be calculated from 
-[International Labor Organization][link-ilo] (ILO) and
-[Food and Agriculture Organization][link-fao] (FAO) data,
-following the methodology of @Steenwyk:2022ww, 
-using functions in the [MWTools][link-mwtools] package [@Marshall:2023ab]. 
-The muscle work data must also be converted to RUVY matrices.
-Third, the IEA's primary- and final-stage ECC data 
-must be extended to the useful stage by
-(a) allocating final stage energy to end-use machines and
-(b) multiplying allocated final energy by the 
-final-to-useful efficiency of each machine.
-This task is accomplished by functions in the
-[Recca][link-recca] [@Heun-Recca:2023]
-package.
-Fourth, ECCs must be converted from energy terms to exergy terms. 
-This step is also performed by functions in the [Recca][link-recca] package.
-
-The steps to create the PSUT matrices for each country and each year
-are accomplished by a
-[targets][link-targets] [@Landau:2021aa]
-computation pipeline
-available in the [PFUDatabase][link-pfudatabase]
-package.
-A unique feature of the
-[PFUDatabase][link-pfudatabase]
-pipeline enabled by the [Recca][link-recca] package 
-is an innovative exemplar system that allows
-analyses to proceed when allocation or efficiency data are unavailable for a country.
-When allocations or efficiencies are missing,
-a string of exemplar countries or regions
-are queried for the required information.
-For example, exemplars for Belgium are
-France, Europe, and ultimately, the World, in that order.
-The design of the calculation pipeline in the [PFUDatabase][link-pfudatabase] package 
-allows allocation and efficiency data
-for any country or year to be added at a later date 
-to improve the database.
-
-A second [targets][link-targets] pipeline 
-in the [PFUAggDatabase][link-pfuaggdatabase] package [@Heun-PFUAggDatabase:2023]
-aggregates ECCs
-(a) by region 
-(e.g., continents and world),
-(b) by energy carrier (product) category 
-(e.g., Coal and coal products, 
-Low-, Medium-, and High-temperature heat, etc.), 
-(c) by energy conversion machines (industries) and/or end use sectors
-(e.g., Residential, Transport, etc.), and
-(d) to primary, final, and useful (PFU) stages.
-In addition, the [PFUAggDatabase][link-pfuaggdatabase] pipeline 
-calculates efficiencies for each country
-at all available aggregations.
-The [PFUAggDatabase][link-pfuaggdatabase]
-pipeline uses the [Recca][link-recca] package extensively.
-
-Both [targets][link-targets] pipelines 
-(in the [PFUDatabase][link-pfudatabase] and
-[PFUAggDatabase][link-pfuaggdatabase] packages)
-produce data frames readable by the [pins][link-pins] package [@Silga:2023].
-Both pipelines benefit from the 
-[PFUSetup][link-pfusetup] [@Heun-PFUSetup:2023] and
-[PFUPipelineTools][link-pfupipelinetools] [@Heun-PFUPipelineTools:2023]
-packages. 
-[PFUSetup][link-pfusetup] identifies storage locations 
-for pipeline input and output data.
-[PFUPipelineTools][link-pfupipelinetools] provides functions and constants
-common to both pipelines.
-
+Several additional packages support the creation of the CL-PFU database. 
 The packages in the following table are available on GitHub.
@@
 
-Input data for the PFU database can be found in @Marshall:2023aa.
-Access to the full CL-PFU database can be obtained via correspondence 
-with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the PFU Database contains
-primary and final energy [IEA EWEB][link-ieaeweb] data, 
-use of the PFU Database
-is restricted to those who have access to [IEA][link-iea] [EWEB][link-ieaeweb] data.].
-If desired, researchers with access to the 
-[IEA][link-iea]'s [EWEB][link-ieaeweb] 
-data can create their own CL-PFU database 
-following the example below.
-
-# Example
-
-```{r, eval=FALSE}
-# (1) Purchase, download, and store IEA EWEB data in the correct location.
-#     Sample data for two countries (Ghana and South Africa) 
-#     and two years (1971 and 2000) are included (with permission)
-#     at the correct location in the repository for this paper.
-# (2) Download FAO and ILO data for muscle work calculations
-#     via the included script.
-#     This download step is needed only once, and
-#     downloaded data are stored in the repository 
-#     for this paper.
-source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
-```
-
-```{r, eval=FALSE}
-# (3) Run the PFUDatabase pipeline to create RUVY matrices.
-targets::tar_make_future(
-  script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
-  workers = 4
-)
-# (4) Run the PFUAggDatabase pipeline to aggregate 
-#     energy and calculate efficiencies.
-targets::tar_make_future(
-  script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
-  workers = 4
-)
-```
-
-```{r}
-# (5) Review results using the pins package.
-#     Establish the pinboard.
-pinboard <- file.path("ExampleFolder", "OutputData", "PipelineReleases") |> 
-  pins::board_folder(versioned = TRUE)
-#     Look at the psut data frame
-#     created by the PFUDatabase pipeline.
-psut <- pinboard |>
-  pins::pin_read(name = "psut", version = "20230811T122618Z-3b7b3")
-head(as.data.frame(psut), 2)
-#     Each column of the psut data frame contains sparse matrices, 
-#     the columns of which are suppressed on display.
-#     Show the column names here.
-colnames(psut$R[[1]])
-#     Show an example resource (R) matrix.
-psut$R[[1]]
-
-#     Review the aggregate PFU efficiency data frame
-#     created by the PFUAggDatabase pipeline.
-agg_eta_pfu <- pinboard |> 
-  pins::pin_read(name = "agg_eta_pfu", version = "20230817T180346Z-45741")
-agg_eta_pfu |> 
-#     Filter and delete columns 
-#     to present a subset of exergy results in a data frame.
-#     The EX.p, EX.f, and EX.u columns contain aggregate exergy at
-#     primary, final, and useful stages, respectively, in TJ.
-#     The eta_pf, eta_fu, and eta_pu columns contain aggregate
-#     primary-to-final, final-to-useful, and 
-#     primary-to-useful efficiencies.
-  dplyr::filter(Energy.type == "X", IEAMW == "Both", 
-                Product.aggregation == "Specified", 
-                Industry.aggregation == "Specified", 
-                GrossNet == "Net",
-                Country %in% c("GHA", "ZAF")) |> 
-  dplyr::mutate(
-    Method = NULL, 
-    Energy.type = NULL,
-    Product.aggregation = NULL,
-    Industry.aggregation = NULL,
-    IEAMW = NULL,
-    Chopped.mat = NULL, 
-    Chopped.var = NULL, 
-    GrossNet = NULL
-  )
-```
-
-# Conclusion
-
-The new open source `R` packages described in this paper
-enable, for the first time, 
-creation of databases of primary, final, and useful
-energy and exergy conversion chains
-in a scalable manner.
-Such databases will enable advances in societal exergy analysis
-and lead to new insights 
-about the role of energy 
-in economic growth and human well-being.
-An example of early work using the CL-PFU database
-is the calculation of useful stage 
-Energy Return On Investment (EROI) for fossil fuels [@Aramendia:2023aa], 
-for which the [ECCTools][link-ecctools] [@Aramendia:2022aa] and
-[EROITools][link-eroitools] [@Aramendia:2022ab] packages were developed, 
-expanding the software capabilities introduced here^[Note
-that the [ECCTools][link-ecctools] and [EROITools][link-eroitools] packages 
-are not the subject of this paper and were not peer reviewed.].
-
-# Acknowledgements
-
-MKH acknowledges the support of Calvin University for 
-a sabbatical during which he developed the PSUT framework
-and course releases during which he wrote many of the `R` packages
-described in this paper.
-ZM's and PEB's time was funded by 
-the UK Research Council under EPSRC Fellowship award EP/R024254/1.
-EA's time was funded by 
-the University of Leeds (School of Earth and Environment).
-
-# References
+## References

### 2023-12-11 paper.Rmd
Addressing #12

@@
 we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
-to assist.
+to support the PSUT framework.
 The table below summarizes these packages, 
@@
 
-The packages in the following table are available on GitHub.
+The packages in the following table support creation of the CL-PFU database
+are available on GitHub.
 

### 2023-12-11 paper.Rmd
Updating example

@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
 # (1) Purchase, download, and store IEA EWEB data in the correct location.
@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
 # (3) Run the PFUDatabase pipeline to create RUVY matrices.

### 2023-12-11 ExampleFolder/_targets_pfudatabase.R

@@ -16,6 +16,8 @@ pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
 # Custom parameters
 countries <- c("GHA", "ZAF")
 years <- c(1971, 2000)  
+years_exiobase <- 1995:2020
+
 
 # Additional exemplar countries are countries which aren't included in the workflow
 # as individual countries, but from which allocation or efficiency data may be 
@@ -89,6 +91,8 @@ PFUDatabase::get_pipeline(countries = countries,
                           machine_data_path = pfu_setup[["machine_data_folder"]],
                           exemplar_table_path = pfu_setup[["exemplar_table_path"]],
                           fu_analysis_folder = pfu_setup[["fu_analysis_folder"]],
+                          exiobase_energy_flows_path = pfu_setup[["exiobase_energy_flows_path"]],
+                          years_exiobase = years_exiobase,
                           reports_source_folders = pfu_setup[["reports_source_folders"]],
                           reports_dest_folder = pfu_setup[["reports_dest_folder"]],
                           pipeline_releases_folder = pfu_setup[["pipeline_releases_folder"]],

### 2023-12-11 README.Rmd
Update README with statement of need.

@@
 output: github_document
-bibliography: References.bib
+bibliography: paper.bib
 ---
@@
 covering the suite of `R` packages associated with the 
-CL-PFU database^["CL-PFU" 
+CL-PFU^["CL-PFU" 
 is an initialism for "Country-Level
@@
 stored as [pins][link-pins] on a pinboard, 
-not SQL or similar databases.].
+not SQL or similar databases.] database.
+
+## Statement of need
+
+Historically, Societal Exergy Analysis (SEA)
+practitioners have analyzed the ECCs
+of individual countries using linked spreadsheets,
+often starting with primary- and final-stage data 
+from the [International Energy Agency][link-iea]'s (IEA's)
+[extended world energy balances][link-ieaeweb] (EWEBs)
+[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
+Data were stored in varying and inconsistent formats.
+An early SEA database [@De-Stercke:2014]
+estimated energy efficiencies 
+of end-use machines using the economic status of countries,
+thereby precluding use in energy-economy studies.
+
+The authors of the current paper and others in the field
+wanted to expand SEA to cover all countries, but
+the spreadsheet approach to SEA was
+deemed not scalable.
+A new approach to SEA was needed, 
+one that scaled across all countries 
+for many years without relying on economic data
+to estimate machine efficiencies.
 
@@
 The CL-PFU database is a new resouce 
-for the Societal Exergy Analysis (SEA) research community and
+for the SEA research community and
 assists SEA practitioners 

### 2023-12-11 paper.Rmd
Working on example

@@
 source(file.path("ExampleFolder", "DownloadScripts", "download_mw_data.R"))
-```
 
-```{r, eval=TRUE}
+# Steps (3) and (4) to construct the database are needed only once.
+# Results are stored in the targets cache 
+# located in the "_targets" folder
+# at the root level of this repository.
+
 # (3) Run the PFUDatabase pipeline to create RUVY matrices.
@@
 psut <- pinboard |>
-  pins::pin_read(name = "psut", version = "20230811T122618Z-3b7b3")
+  pins::pin_read(name = "psut", version = "20231211T161342Z-3e2fb")
 head(as.data.frame(psut), 2)
@@
 agg_eta_pfu <- pinboard |> 
-  pins::pin_read(name = "agg_eta_pfu", version = "20230817T180346Z-45741")
+  pins::pin_read(name = "agg_eta_pfu", version = "20231211T161827Z-8f5e7")
 agg_eta_pfu |> 

### 2023-12-11 ExampleFolder/_targets_pfudatabase.R

@@ -44,7 +44,7 @@ apply_fixes <- TRUE
 
 # Should we do a release of the results to a pinboard?
 # A release has already been made and is available in the repository.
-release <- FALSE
+release <- TRUE
 
 # End user-adjustable parameters.
 

### 2023-12-12 paper.Rmd
Paper now able to build again.
Added metapackage information

--- 2023-12-11 paper.Rmd
@@
 We created a suite of open-source `R` packages 
+and the metapackage `CLPFUDatabase`
 to assist SEA practitioners to analyze ECCs. 
@@
 we developed the [RCLabels][link-rclabels] package [@Heun-RCLabels:2023]
-to support the PSUT framework.
+to assist.
 The table below summarizes these packages, 
@@
 
-The packages in the following table support creation of the CL-PFU database
-are available on GitHub.
+The packages in the following table are available on GitHub.
 

### 2023-12-12 paper.Rmd

--- 2023-12-11 paper.Rmd
@@
 
-```{r, eval=TRUE}
+```{r, eval=FALSE}
 # (1) Purchase, download, and store IEA EWEB data in the correct location.

### 2023-12-12 paper.Rmd
Add metapackage to table.

@@
 [link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
+[link-clpfudatabase]:    https://github.com/EnergyEconomyDecoupling/CLPFUDatabase
 [link-pins]:             https://pins.rstudio.com
@@
 | [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
+| [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all packages for the CL-PFU database |
 
@@
 Access to the full CL-PFU database can be obtained via correspondence 
-with author [PEB](mailto:P.E.Brockway@leeds.ac.uk)^[Because the CL-PFU database contains
+with author [P.E.B.](mailto:P.E.Brockway@leeds.ac.uk)^[Because the CL-PFU database contains
 primary and final energy [IEA EWEB][link-ieaeweb] data, 

### 2023-12-18 paper.Rmd
Remove PEB from author list.

@@
   orcid: 0000-0001-9260-7827
-  affiliation: 2
-- name: Paul E. Brockway
-  orcid: 0000-0001-6925-8040
   affiliation: 2

### 2023-12-18 paper.Rmd
Add Acknowledgement for PEB.

@@
 
-MKH acknowledges the support of Calvin University for 
+M.K.H. acknowledges the support of Calvin University for 
 a sabbatical during which he developed the PSUT framework
@@
 described in this paper.
-ZM's and PEB's time was funded by 
+Z.M.'s time was funded by 
 the UK Research Council under EPSRC Fellowship award EP/R024254/1.
-EA's time was funded by 
+E.A.'s time was funded by 
 the University of Leeds (School of Earth and Environment).
+M.K.H., Z.M., and E.A. are grateful for support and encouragement
+from Paul E. Brockway (University of Leeds).
+It is Paul's EPSRC fellowship that provided the need for the software
+described in this paper.
 

### 2023-12-19 paper.Rmd
Updated acknowledgements.
Updated paper.Rmd

@@
 | [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
-| [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all packages for the CL-PFU database |
+| [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all package dependencies for the CL-PFU database |
 
@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
+library(devtools)
+# (0) Use the metapackage to install all dependencies.
+#     This package installation step is needed only once.
+devtools::install_github("EnergyEconomyDecoupling/CLPFUDatabase")
+
 # (1) Purchase, download, and store IEA EWEB data in the correct location.
@@
 M.K.H., Z.M., and E.A. are grateful for support and encouragement
-from Paul E. Brockway (University of Leeds).
-It is Paul's EPSRC fellowship that provided the need for the software
+from Dr. Paul E. Brockway (University of Leeds).
+It was Dr. Brockway's EPSRC fellowship that provided the need for the software
 described in this paper.

### 2023-12-19 paper.Rmd
Better comments.
Completed the run

@@
 
-```{r, eval=TRUE}
+```{r, eval=FALSE}
 library(devtools)
-# (0) Use the metapackage to install all dependencies.
+# (0) Use the CLPFUDatabase metapackage to install all dependencies.
 #     This package installation step is needed only once.
@@
 psut <- pinboard |>
-  pins::pin_read(name = "psut", version = "20231211T161342Z-3e2fb")
+  pins::pin_read(name = "psut", version = "20231219T204403Z-2c3ce")
 head(as.data.frame(psut), 2)

### 2023-12-19 ExampleFolder/_targets_pfuaggdatabase.R

@@ -19,8 +19,8 @@ do_chops <- FALSE
 
 # Set the release to be used for input.
 # This is a pre-made pins release available in the repository.
-psut_release <- "20231211T161342Z-3e2fb"  
-psut_without_neu_release <- "20231211T161342Z-bf294"
+psut_release <- "20231219T204403Z-2c3ce"  
+psut_without_neu_release <- "20231219T204408Z-6761d"
 phi_vecs_release <- "20231211T161340Z-78875"
 
 # Should we release the results?

### 2023-12-20 paper.Rmd
Add empty line.
Eliminate date field.

@@
   index: 2
-# date: "17 Aug 2023"
 bibliography: paper.bib
@@
 #     at the correct location in the repository for this paper.
+
 # (2) Download FAO and ILO data for muscle work calculations

### 2023-12-20 paper.Rmd
Add a blank line.

@@
 )
+
 # (4) Run the PFUAggDatabase pipeline to aggregate 

### 2023-12-20 paper.Rmd
Added ERE paper reference.

@@
 not SQL or similar databases.],
-a new resource for the SEA community [@Marshall:2023aa].
+a new resource for the SEA community [@Marshall:2023aa; @Brockway:2024aa].
 

### 2023-12-20 README.Rmd
Added metapackage information
PFUDatabase-->PFUPipeline; PFUAggDatabase-->PFUAggPipeline

@@
 [link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
-[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
-[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
+[link-pfupipeline]:      https://github.com/EnergyEconomyDecoupling/PFUPipeline
+[link-pfuaggpipeline]:   https://github.com/EnergyEconomyDecoupling/PFUAggPipeline
 [link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
@@
 We created a suite of `R` packages 
+and the metapackage `CLPFUDatabase`
 to support the creation and use of the CL-PFU database.
@@
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
-| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
+| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUPipeline][link-pfupipeline] and [PFUAggPipeline][link-pfuaggpipeline] pipelines |
 | [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
-| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
-| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
+| [PFUPipeline][link-pfupipeline]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
+| [PFUAggPipeline][link-pfuaggpipeline] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
 

### 2023-12-20 paper.Rmd
PFUDatabase-->PFUPipeline; PFUAggDatabase-->PFUAggPipeline

@@
 computation pipeline
-available in the [PFUDatabase][link-pfudatabase]
+available in the [PFUPipeline][link-pfupipeline]
 package.
 A unique feature of the
-[PFUDatabase][link-pfudatabase]
+[PFUPipeline][link-pfupipeline]
 pipeline enabled by the [Recca][link-recca] package 
@@
 France, Europe, and ultimately, the World, in that order.
-The design of the calculation pipeline in the [PFUDatabase][link-pfudatabase] package 
+The design of the calculation pipeline in the [PFUPipeline][link-pfupipeline] package 
 allows allocation and efficiency data
@@
 A second [targets][link-targets] pipeline 
-in the [PFUAggDatabase][link-pfuaggdatabase] package [@Heun-PFUAggDatabase:2023]
+in the [PFUAggPipeline][link-pfuaggpipeline] package [@Heun-PFUAggPipeline:2023]
 aggregates ECCs
@@
 Both [targets][link-targets] pipelines 
-(in the [PFUDatabase][link-pfudatabase] and
-[PFUAggDatabase][link-pfuaggdatabase] packages)
+(in the [PFUPipeline][link-pfupipeline] and
+[PFUAggPipeline][link-pfuaggpipeline] packages)
 produce data frames readable by the [pins][link-pins] package [@Silga:2023].
@@
 | [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
-| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
+| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUPipeline][link-pfupipeline] and [PFUAggPipeline][link-pfuaggpipeline] pipelines |
 | [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
-| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
-| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
+| [PFUPipeline][link-pfupipeline]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
+| [PFUAggPipeline][link-pfuaggpipeline] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
 | [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all package dependencies for the CL-PFU database |
@@
 
-# (3) Run the PFUDatabase pipeline to create RUVY matrices.
+# (3) Run the PFUPipeline to create RUVY matrices.
 targets::tar_make_future(
-  script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
+  script = file.path("ExampleFolder", "_targets_pfupipeline.R"), 
   workers = 4
@@
 
-# (4) Run the PFUAggDatabase pipeline to aggregate 
+# (4) Run the PFUAggPipeline to aggregate 
 #     energy and calculate efficiencies.
 targets::tar_make_future(
-  script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
+  script = file.path("ExampleFolder", "_targets_pfuaggpipeline.R"),
   workers = 4
@@
 #     Look at the psut data frame
-#     created by the PFUDatabase pipeline.
+#     created by the PFUPipeline.
 psut <- pinboard |>
@@
 #     Review the aggregate PFU efficiency data frame
-#     created by the PFUAggDatabase pipeline.
+#     created by the PFUAggPipeline.
 agg_eta_pfu <- pinboard |> 

### 2023-12-20 ExampleFolder/_targets_pfuaggpipeline.R

@@ -2,7 +2,7 @@ library(targets)
 
 # Use the PFUSetup package to establish paths for the example
 # Create the pfu_setup object used in the targets scripts 
-# _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
+# _targets_pfupipeline.R and _targets_pfuaggpipeline.R.
 pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
                                      project_path = "ExampleFolder",
                                      version = "v_example", 
@@ -35,7 +35,7 @@ future::plan(future.callr::callr)
 
 # Set options for all targets.
 targets::tar_option_set(
-  packages = "PFUAggDatabase",
+  packages = "PFUAggPipeline",
   # Indicate that storage and retrieval of subtargets
   # should be done by the worker thread,
   # not the main thread.

### 2023-12-20 ExampleFolder/_targets_pfupipeline.R

@@ -2,12 +2,12 @@
 
 # Available from EnergyEconomyDecoupling on GitHub
 # with devtools::install_github("EnergyEconomyDecoupling/xxxxxx")
-library(PFUDatabase)
+library(PFUPipeline)
 library(PFUSetup)
 
 # Use the PFUSetup package to establish paths for the example
 # Create the pfu_setup object used in the targets scripts 
-# _targets_pfudatabase.R and _targets_pfuaggdatabase.R.
+# _targets_pfupipeline.R and _targets_pfuaggpipeline.R.
 pfu_setup <- PFUSetup::get_abs_paths(home_path = "", cloud_storage_path = "",
                                      project_path = "ExampleFolder",
                                      version = "v_example", 
@@ -70,7 +70,7 @@ targets::tar_option_set(
 )
 
 # Pull in the pipeline
-PFUDatabase::get_pipeline(countries = countries,
+PFUPipeline::get_pipeline(countries = countries,
                           additional_exemplar_countries = additional_exemplar_countries,
                           matrix_class = matrix_class,
                           specify_non_energy_flows = specify_non_energy_flows,

### 2023-12-20 paper.Rmd
AggDatabase --> AggPipeline

@@
 [PFUPipeline][link-pfupipeline]
-pipeline enabled by the [Recca][link-recca] package 
+enabled by the [Recca][link-recca] package 
 is an innovative exemplar system that allows
@@
 (d) to primary, final, and useful (PFU) stages.
-In addition, the [PFUAggDatabase][link-pfuaggdatabase] pipeline 
+In addition, the [PFUAggPipeline][link-pfuaggpipeline] 
 calculates efficiencies for each country
 at all available aggregations.
-The [PFUAggDatabase][link-pfuaggdatabase]
-pipeline uses the [Recca][link-recca] package extensively.
+The [PFUAggPipeline][link-pfuaggpipeline]
+uses the [Recca][link-recca] package extensively.
 

### 2023-12-20 ExampleFolder/_targets_pfuaggpipeline.R

@@ -46,7 +46,7 @@ targets::tar_option_set(
 )
 
 # Pull in the pipeline
-PFUAggDatabase::get_pipeline(countries = countries,
+PFUAggPipeline::get_pipeline(countries = countries,
                              years = years,
                              do_chops = do_chops,
                              psut_release = psut_release,

### 2023-12-21 paper.Rmd
Remove library(devtools)
Update comments.

@@
 ```{r, eval=FALSE}
-library(devtools)
 # (0) Use the CLPFUDatabase metapackage to install all dependencies.
-#     This package installation step is needed only once.
+#     This metapackage installation step is needed only once.
 devtools::install_github("EnergyEconomyDecoupling/CLPFUDatabase")

### 2023-12-21 paper.Rmd
Added link and reference to the CLPFUDatabase package

@@
 We created a suite of open-source `R` packages 
-and the metapackage `CLPFUDatabase`
+and the metapackage [CLPFUDatabase][link-clpfudatabase] [@Heun-CLPFUDatabase:2023]
 to assist SEA practitioners to analyze ECCs. 
@@
 We used the new packages to create the CL-PFU database^["CL-PFU" 
-is an initialism for "Country-Level
-Primary, Final, and Useful."
+is an initialism for 
+"Country-Level Primary, Final, and Useful."
 The CL-PFU database contains many countries as well as 

### 2023-12-21 README.Rmd
Added CLPFUDatabase to table in README.Rmd to solve #15

@@
 | [PFUAggPipeline][link-pfuaggpipeline] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
+| [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all package dependencies for the CL-PFU database |
 

### 2024-01-03 paper.Rmd
Quoting affiliations

@@
   orcid: 0000-0002-7438-214X
-  affiliation: 1,2
+  affiliation: "1,2"
 - name: Zeke Marshall

### 2024-01-03 paper.Rmd
Add 3rd affiliation for MKH

--- 2023-12-21 paper.Rmd
@@
   orcid: 0000-0002-7438-214X
-  affiliation: 1,2
+  affiliation: "1,2,3"
 - name: Zeke Marshall
@@
   index: 2
+- name: School for Public Leadership, Stellenbosch University
+  index: 3
 bibliography: paper.bib

### 2024-01-08 paper.Rmd
PFU database --> CL-PFU database
Remove references to repository DOI.

@@
 not SQL or similar databases.],
-a new resource for the SEA community [@Marshall:2023aa; @Brockway:2024aa].
+a new resource for the SEA community [@Brockway:2024aa].
 
@@
 
-Input data for the PFU database can be found in @Marshall:2023aa.
+Example input data for the CL-PFU database 
+(for two countries and two years) 
+can be found in the GitHub repository for this paper.
 Access to the full CL-PFU database can be obtained via correspondence 

### 2024-01-08 paper.Rmd
Minor changes.

@@
 This task is accomplished by functions in the
-[Recca][link-recca] [@Heun-Recca:2023]
-package.
+[Recca][link-recca]
+package [@Heun-Recca:2023].
 Fourth, ECCs must be converted from energy terms to exergy terms. 
@@
 available in the [PFUPipeline][link-pfupipeline]
-package.
+package [@Heun-PFUPipeline:2023].
 A unique feature of the
@@
 
-The packages in the following table support creation of the CL-PFU database
+The packages in the following table support creation of the CL-PFU database and
 are available on GitHub.

### 2024-01-09 paper.Rmd
Capitalize EWEB

@@
 from the [International Energy Agency][link-iea]'s (IEA's)
-[extended world energy balances][link-ieaeweb] (EWEBs)
+[Extended World Energy Balances][link-ieaeweb] (EWEBs)
 [@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 

### 2024-01-09 README.Rmd
Add JOSS badge to README

@@
 
+[![DOI](https://joss.theoj.org/papers/10.21105/joss.06057/status.svg)](https://doi.org/10.21105/joss.06057)
+
 ## JOSS paper
 
-This repository contains a paper submitted to 
+This repository contains a paper published in  
 the [Journal of Open Source Software (JOSS)](https://joss.theoj.org)
